### 2012-10-26 paper.Rnw
Initial commit after everything died because dropbox was being cool.

\documentclass{article}
\usepackage{spconf,amsmath,graphicx}

\newcommand{\todo}[1]{
    \addcontentsline{tdo}{todo}{\protect{#1}}
    \textcolor{red}{#1}
}

\title{HIPPOCAMPAL SEGMENTATION}
\name{  Jon Pipitone$^{\star}$ Jason P. Lerch$^{\dagger}$ Miriam Friedel$^{\dagger}$ Aristotle Voineskos$^{\star}$
         Mallar Chakravarty$^{\star}$ }

\address{$^{\star}$ Centre for Additions and Mental Health, Toronto ON, Canada \\
     $^{\dagger}$ Mouse Imaging Centre, ..., Toronto ON, Canada}

\begin{document}
\SweaveOpts{concordance=TRUE}
\maketitle

\begin{abstract}
Neuroimaging research often relies on automated anatomical
segmentations of MR images of the brain. Current multi-atlased based
approaches provide accurate segmentations of brain images by
propagating region labels from manual delineations to unlabeled
images.  Unfortunately, these approaches often rely on a large number
of such manual segmentations which take time and significant expertise
to produce, neither of which may be readily available.  We present
an algorithm for the automatic segmentation of the hippocampus that
minimizes the number of atlases needed whilst still achieving nearly
the same accuracy as other multi-atlas approaches.  We perform
repeated random sub-sampling validation on the IBSR dataset to
validate our approach and find an average best case percent difference
to the multi-atlas approach of 1.9\%.
% fix last line: clarify "best case percent difference"

\end{abstract}

\begin{keywords}
hippocampus, segmentation, multi-atlas, automated
\end{keywords}

\section{Introduction}
\label{sec:intro}

The hippocampus is of particular interest to many researchers because it is
implicated in forms of brain dysfunction such as Alzheimer's disease and
schizophrenia, and has functional significance in cognitive processes such as
learning and memory.  For many research questions involving neuroimaging data,
accurate identification of the hippocampus and its subregions in participant MR
images is a necessary first step in order to analyse neuroanatomical
differences and changes in sample populations.

Currently, the gold standard for neuroanatomical segmentation is manual
labelling by an expert human rater.  This is problematic for segmentation of
the hippocampus for several reasons.  First, manual segmentation takes a
significant investment of time and expertise \cite{Hammers2003} which may not
be readily available to researchers or clinicians.  Second, the amount of data
produced in neuroimaging experiments increasingly exceeds the capacity for
identification of specific neuroanatomical structures by an expert manual
rater.  Third, the true delination of hippocampal anatomy in MR images is
disputed\cite{Geuze2004}, as evidenced by efforts to create an unified
segmentation protocol\cite{Jack2011}.  Compounding each of these problems is
that significant neuroanatomical variability in the hippocampus throughout the
development and aging process and across the time course of different diseases
is expected. As well, it may be necessary to use several different hippocampal
definitions or, in fact, make research-specific modifications in order to test
hypotheses (for example, \cite{Poppenk2011} found that overall hippocampal
volume difference did not predict recollection memory performance but by
dividing the hippocampus into anterior and posterior regions, a predictive
volume difference was found).  Thus, whilst manual segmentation of the
hippocampus is an important technique, it may be a bottleneck for researchers
or clinicians who do not have access to the needed human expertise.

Automated segmentation techniques attempt to address the need for human
expertise by performing segmentations computationally.  A popular class of
automated methods, {\it atlas-based segmentation}, rely on a set of expertly
labeled neuroanatomical atlases.  Each atlas is warped to fit a subject's
neuroanatomy using nonlinear registration
techniques\cite{Collins1995,Klein2009}.  Atlas labels are then transformed by
this warping and a {\it label fusion} technique, such as voxel-wise voting, is
used to merge the labellings from each atlas into a final segmentation for a
subject.

Many descriptions of atlas-based segmentation algorithms report relying on an
atlas library containing between 30 and 80 expertly labeled
brains\cite{Heckemann2011,Collins2010,Aljabar2009,Leung2010,Lotjonen2010}.
As noted, the production of an atlas library requires significant manual
effort, and is limited since the choice of atlases or segmentation protocol may
not reflect the underlying neuroanatomical variability of the population under
study or be suited to answer the research questions at hand.

In this paper we propose an automated segmentation technique to address the
issues found in existing atlas-based methods. Principly, our method aims to
dramatically reduce the number of manually labelled atlases necessary.  It does
this by using a small atlas set to {\it generate} a much larger ``template
library'', which is then used to segment the subjects in the same fashion as
other atlas-based methods.  The essential insight of generating a template
library is not new.  Heckemann \cite{Heckemann2006} compared generating a
template library from a single atlas to standard multi-atlas segmentation and
found poor performance and so deemed the approach as inviable.  The LEAP
algorithm \cite{Wolz2010} proceeds by iteratively segmenting the unlabelled
image most similar to the atlas library images and then incorporating the
now-labelled image into the atlas library, but requires 30 starting atlases.
Our contribution is provide a method that provides comparable segmentation
accuracy to these and other atlas-based maethods whilst using significantly
fewer manually created atlases.

In previous work from our group \cite{Chakravarty2011}, we explored
this same approach to the segmentation of certain subcortical structures (striatum,
globus pallidus, and thalamus) using a single histologically-derived atlas.
In this work we extend our approach to the hippocampus, and validate it by
varying the atlas and template library sizes.  Due to the small number of
atlases required, our method could easily accommodate different hippocampal
definitions.  Our aim is not to improve on segmentation accuracy beyond
existing methods, but instead to provide a method that trades off manual
segmentation expertise for computational processing time whilst providing
sufficient accuracy for clinical and research applications.

\section{Materials and Methods} 
\subsection{The Multiple Automatically Generated Templates (MAGeT) Algorithm}

In this paper, we use the term {\it atlas} to mean any manually segmented MR
image, and the term {\it atlas library} to mean a set of such images.  We use
the term {\it template} to refer to any MR image, and associated labelling,
used to segment another image, and the term {\it template library} to refer to
a set of such images.  An atlas library may be used as a template library but,
as we will discuss, a template library may also be composed of images with
computer generated labellings.

The segmentation approach we propose is best understood as an extension of
traditional multi-atlas segmentation.  In multi-atlas segmentation, the inputs
are an atlas library and a set of unlabelled MR images.  An unlabelled image is
segmented in the following way.  Each atlas image is non-linearly registered to
the unlabelled image, and then each atlas' labels are propagated via the
resulting transformation.  These candidate labels are then fused to produce a
definitive segmentation. The label fusion method used may determine which
templates are used to segment a particular subject (for instance, voxel-wise
majority vote takes labels from all atlases, whereas a weighted majority vote
selects only those templates most similar to the subject).

Our extension to this approach is to add a preliminary stage in which we
construct a template library to be used in the standard atlas-based approach as
if it were the atlas library.  To do this we select unlabelled images from the
input images to form the template library.  Next, labels from each atlas image
are propagated to each image in the template library (via the transformation
resulting from a non-linear registration between each pair of images).
Thus, each template library image has a labelling from each atlas.  Traditional
multi-atlas segmentation is then used to produce segmentations for the entire
set of unlabelled images (including those images used in the template library).

We use the {\it cross-correlation weighted voting} method to fuse the
candidate labels of each unlabeled image.  In this method, a voxel-wise
majority vote is carried out the labels from the top $n$ ranked template
library images.  The scores are given by the normalized cross-correlation
between the unlabeled image and each template library image, after the
unlabeled image is linearly registered to the template image.  The
cross-correlation score itself is calculated over a region of interest
(ROI) defined on each template.  A customized ROI is defined by first
linearly registering each unlabeled subject to all possible templates in
the template library, transforming all template hippocampus labels,
defining the possible extent of all hippocampal labels, and finally
dilating this extent by three voxels.
\todo{jp: it isn't clear that only the top ranked images are being used.
perhaps some rationale for CC voting would help here. i.e. discuss improvement
in overall registration quality --  throwing out atlases liable to introduce
unwanted variance }

\subsection{Image Processing and Registration Method}
The N3 algorithm \cite{Sled1998} is first used to minimize the intensity
nonuniformity in each of the atlases and unlabeled subject images.  Image
registration is carried out in two phases.  In the first, a 12-parameter linear
transformation (3 translations, rotations, scales, shears) is estimated between
images using an algorithm that maximizes the correlation between blurred MR
intensities and gradient magnitude over the whole brain \cite{Collins}.  In the
second phase, nonlinear registration is completed using the ANIMAL algorithm
\cite{Collins1995}: an iterative procedure that estimates a 3D deformation
field between two MR images. At first, large deformations are estimated using
blurred version of the input data. These larger deformations are then input to
subsequent steps where the fit is refined by estimating smaller deformations on
data blurred with a Gaussian kernel with a smaller FWHM. The final
transformation is a set of local translations defined on a bed of equally
spaced nodes that were estimated through the optimization of the correlation
coefficient. For the purposes of this work we used the regularization
parameters optimized in Robbins et al.\cite{Robbins2004}.

\subsection{MRI dataset evaluated}

For evaluation purposes we used the publicly available IBSR dataset.  This
dataset consists of T1 weighted MR image volumes from 18 subjects (4 females,
14 males) with ages between 7 and 71 years. Image dimensions for all MR volumes
are normalized to  $256  \times 256 \times 128$ voxels, with the voxel size
ranging from $0.8 \times 0.8 \times 1.5 mm$ to $1.0 \times 1.0 \times 1.5 mm$.
The images come 'positionally normalized' into the Talairach orientation
(rotation only), and processed by the CMA 'autoseg' biasfield correction
routines. The MR brain data sets and their manual segmentations are publicly
available and were provided by the Center for Morphometric Analysis at
Massachusetts General Hospital and are available at
\url{http://www.cma.mgh.harvard.edu/ibsr/}.

\subsection{Experiments}

In our experiments we verified how varying two different parameters effects
labeling accuracy.

\subsubsection{Varying Number of Atlases.}
First we varied the number of input atlases that could be used from IBSR
dataset from 3 to 8.  In each case, we conducted 30 rounds of random
subsampling cross-validation using the 18 manual segmented templates from
the IBSR dataset as input. In each round, atlases were randomly chosen from
the IBSR dataset and the remaining templates are used both as template
library images and unlabeled subjects to be labeled using the MAGeT Brain
algorithm.

\subsubsection{Varying Number of Templates and Cross-Correlation Voting.}
The second parameter we varied was the number of possible candidate
labels to fuse for the final segmentation.  For each round of
validation in Experiment 1, we carried out cross-correlation weighted
label fusion amongst only the labels from the $n$ candidate templates
from the library (i.e. selecting only those labels derived from
templates with highest cross-correlation within the hippocampal ROI).
The number of templates ($n$) was varied in the range $[3, r]$ (where $r$
is the number of possible templates in the library).

%where candidate labels are ranked by the cross-correlation score of the
%subject to the template whence the labels came. We use values of n in the

%In this experiment we compare the performance of MAGeT using three manually
%segmented high-resolution atlases. ??We conduct a single run of the MAGeT
%algorithm on 60 subjects and corresponding hippocampal segmentations from
%the ADNI database. 20 subjects with AD, MCI and normals are used to
%ftemplates from a pool of 60 templates to use in the template library, and
%as target subjects. ??The pool of 60 templates is itself, randomly selected
%from the ADNI database of first screening structural MRIs. ??

\subsection{Evaluation}
\subsubsection{Goodness-of-fit}
Each segmentation was evaluated against the manual segmentation from the IBSR
dataset using the Dice Kappa ($\kappa$) overlap metric:

\[ \kappa = \frac{2a}{2a+b+c} \]

\noindent where $a$ is the number of voxels common to the candidate
segmentation and the gold standard and $b+c$ is the sum of the voxels
uniquely identified by either the automatically generated candidate
segmentation or the gold-standard.

\subsubsection{Comparison Approaches}
The resulting segmentations from each of our experiments is compared to two
alternative segmentation approaches. The {\it naive} approach is that in  which
a single atlas is used to segment an unlabeled subject by directly propagating
labels.  The {\it multi-atlas} approach, as described above, in which we used
cross-correlation to select the top most similar atlases to each subject a
atlas
labels are propagated to an unlabeled subject and the labels are fused using
the cross-correlation voting described above.  In total we evaluated
approximately $52,000$ segmentations for the work presented in this manuscript.

\section{Results}

Sample segmentations from one of the IBSR subjects is shown in Fig.
\ref{montage}.  The segmentations shown are varied across the template
selection for a case where 3 atlases were used.  The figure demonstrates a
reduction in false negatives (voxels labeled by the `gold-standard' only)
in the hippocampal tail, a constant number of false positives (voxels
labeled by the MAGeT Brain method only) in the hippocampal head, and an
increase in the segmentation accuracy as the number of templates used from
the template library are increased.

% TODO: check nomenclature for cross-correlation label fusion (and other
% methods) 
\begin{figure}[h]
\begin{minipage}[b]{1.0\linewidth}
  \centering
  \includegraphics[width=\textwidth]{montage.png}
\end{minipage}
\caption{{\em Sample MAGeT Brain segmentations.} Segmentations of a single
subject are shown when 3 atlases are used, with cross-correlation label
fusion. Blue represents agreement between the gold-standard and MAGeT
Brain.  Red indicates voxels labeled only by the gold-standard, and green
indicates voxels labeled only by MAGeT Brain. Note improved accuracy as the
number of templates is increased.} 
\label{montage}
\end{figure}

In Fig. \ref{results} we compare the performance of the MAGeT Brain method
across all validations (i.e. a combination of all atlases and template
selections) against the multi-atlas method and the naive approach.
Regardless of the number of atlases and templates used in a particular
segmentation we see a marked increase in segmentation accuracy over the
average naive segmentation (the average Kappa of all segmentations derived
from all pairwise registrations in the IBSR dataset).  Clearly, including
more input atlases increases segmentation accuracy.  Surprisingly,
increasing the number of labels in the label fusion step through
cross-correlation based voting, does little to improve segmentation
accuracy (except for in the cases where 3 and 4 atlases are used).
Finally, the best-case MAGeT Brain performance scenario is when 8 atlases
are used.  While this does not reach the accuracy of the multi-atlas
segmentation, there is only a 1.9\% difference in the average Kappa values
($0.775$ for MAGeT Brain with 8 atlases and $0.790$ for multi-atlas after
selecting the top 14 templates).

\begin{figure}
\begin{minipage}[b]{1.0\linewidth}
  \centering
  %\includegraphics[width=\textwidth]{results}
<<fig=TRUE, cached=TRUE, echo=FALSE, warnings=FALSE, >>=
library(ggplot2)
results = read.csv(file='data/results.csv')
xcorr_vote= subset(results, method=='xcorr' | method == 'multiatlas_xcorr_vote' )
qplot(x=top_n, y=k_hippocampus, data=xcorr_vote, colour=as.factor(num_atlases), geom=c('point'), alpha=I(0)) +
  geom_smooth(line=1.5, formula=y~poly(x,2), method='lm') + 
  geom_hline(aes(yintercept = mean(subset(results, method=='naive')$k_hippocampus))) + 
  theme(legend.position="none") + 
  ylab('Kappa') + 
  xlab('Number of Atlases') +
  annotate("text", 16.0, 0.712, label="Naive", size=4, hjust=0)+
  annotate("text", 15.3, 0.754, label="3 atlases", size=4, hjust=0)+
  annotate("text", 14.4, 0.766, label="4 atlases", size=4, hjust=0)+
  annotate("text", 13.5, 0.770, label="5 atlases", size=4, hjust=0)+
  annotate("text", 12.5, 0.773, label="6 atlases", size=4, hjust=0)+
  annotate("text", 11.5, 0.776, label="7 atlases", size=4, hjust=0)+
  annotate("text", 10.3, 0.779, label="8 atlases", size=4, hjust=0)+
  annotate("text", 14.6, 0.793, label="Multi-atlas", size=4, hjust=0)+
  coord_cartesian(ylim=c(0.7,0.8), xlim=c(3,18)) + 
  scale_y_continuous(breaks=seq(0,1,0.01))

@  
\end{minipage}
\caption{{\em Performance of MAGeT Brain across all validations compared to
mean naive segmentation, and mean multi-atlas with cross-correlation voting.}
Multi-atlas and naive segmentations procedures provide upper and lower
bounds (respectively) for accuracy.  Note that improved segmentation
accuracy is achieved through increasing the number of input atlases.}
\label{results}
\end{figure}

\section{Discussion}

In this paper, we have demonstrated that accurate segmentations can be
achieved by simply automatically deriving a template library from a small
set of input atlases.  MAGeT Brain segmentations were compared to both
naive segmentations and multi-atlas segmentations with cross-correlation
voting.  In \cite{Lotjonen2010}, the authors report a Kappa of 0.814 on
hippocampal segmentation in the IBSR dataset using a multi-atlas approach
where atlas selection is based on the similarity between atlas and
unlabeled subject after nonlinear registration, and using STAPLE
\cite{Warfield2004} for label fusion.  We demonstrated that the trade off
in accuracy over a multi-atlas segmentation using the entire IBSR manually
segmented library is only on average 1.9\%, using only 8 randomly chosen
atlases.  Discrepancies between our results and the above results may be
due to choice of registration algorithm, regularization parameters, or
similarity metric for label fusion.  Performance may also be affected by
the variability in the IBSR data set (as previously noted by
\cite{Klein2009}).  Future work from our group will attempt to address some
of these issues.

\todo{discuss CC voting and why accuracy is monotonically increasing...
wouldn't we expect an increase and then decrease?}

\bibliographystyle{IEEEbib}
\bibliography{references}
\end{document}

### 2012-10-27 paper.tex
Updated a little of the methods, and added color/hyperref packages.

--- 2012-10-26 paper.Rnw
@@
 \documentclass{article}
-\usepackage{spconf,amsmath,graphicx}
+\usepackage{spconf,amsmath,graphicx,color,hyperref}
 
 \newcommand{\todo}[1]{
-    \addcontentsline{tdo}{todo}{\protect{#1}}
     \textcolor{red}{#1}
@@
 
-\title{HIPPOCAMPAL SEGMENTATION}
+\title{HIPPOCAMPAL SEGMENTATION WITH A VERY SMALL ATLAS LIBRARY: THE MAGET
+BRAIN ALGORITHM} 
 \name{  Jon Pipitone$^{\star}$ Jason P. Lerch$^{\dagger}$ Miriam Friedel$^{\dagger}$ Aristotle Voineskos$^{\star}$
@@
 
+\usepackage{Sweave}
 \begin{document}
-\SweaveOpts{concordance=TRUE}
+\input{paper-concordance}
 \maketitle
@@
 
-We use the {\it cross-correlation weighted voting} method to fuse the
-candidate labels of each unlabeled image.  In this method, a voxel-wise
-majority vote is carried out the labels from the top $n$ ranked template
-library images.  The scores are given by the normalized cross-correlation
-between the unlabeled image and each template library image, after the
-unlabeled image is linearly registered to the template image.  The
-cross-correlation score itself is calculated over a region of interest
-(ROI) defined on each template.  A customized ROI is defined by first
-linearly registering each unlabeled subject to all possible templates in
-the template library, transforming all template hippocampus labels,
-defining the possible extent of all hippocampal labels, and finally
-dilating this extent by three voxels.
-\todo{jp: it isn't clear that only the top ranked images are being used.
-perhaps some rationale for CC voting would help here. i.e. discuss improvement
-in overall registration quality --  throwing out atlases liable to introduce
-unwanted variance }
+We use the {\it cross-correlation weighted voting} method to both improve the
+quality of the label fusion step and reduce the number of registrations
+required \todo{ref}.  In this method, each template library image is ranked 
+in similarity to each unlabelled image, and only the top $n$ template library
+images labels are used for the voxel-wise majority vote fusion step.
+Similarity scores are computed by the normalized cross-correlation of image
+intensities between an unlabeled image and each template library image, after
+the unlabeled image has undergone linearly registration to the template image.
+The cross-correlation score is calculated over a region of interest (ROI) on
+each template representing an generous approximation of hippocampal structure
+and neighbourhood.  The ROI is defined as the extent of all atlas labels after
+linear registration to the template, dilated by three voxels.
 
@@
 Massachusetts General Hospital and are available at
-\url{http://www.cma.mgh.harvard.edu/ibsr/}.
+\url{''http://www.cma.mgh.harvard.edu/ibsr/''}.
 
@@
   %\includegraphics[width=\textwidth]{results}
-<<fig=TRUE, cached=TRUE, echo=FALSE, warnings=FALSE, >>=
-library(ggplot2)
-results = read.csv(file='data/results.csv')
-xcorr_vote= subset(results, method=='xcorr' | method == 'multiatlas_xcorr_vote' )
-qplot(x=top_n, y=k_hippocampus, data=xcorr_vote, colour=as.factor(num_atlases), geom=c('point'), alpha=I(0)) +
-  geom_smooth(line=1.5, formula=y~poly(x,2), method='lm') + 
-  geom_hline(aes(yintercept = mean(subset(results, method=='naive')$k_hippocampus))) + 
-  theme(legend.position="none") + 
-  ylab('Kappa') + 
-  xlab('Number of Atlases') +
-  annotate("text", 16.0, 0.712, label="Naive", size=4, hjust=0)+
-  annotate("text", 15.3, 0.754, label="3 atlases", size=4, hjust=0)+
-  annotate("text", 14.4, 0.766, label="4 atlases", size=4, hjust=0)+
-  annotate("text", 13.5, 0.770, label="5 atlases", size=4, hjust=0)+
-  annotate("text", 12.5, 0.773, label="6 atlases", size=4, hjust=0)+
-  annotate("text", 11.5, 0.776, label="7 atlases", size=4, hjust=0)+
-  annotate("text", 10.3, 0.779, label="8 atlases", size=4, hjust=0)+
-  annotate("text", 14.6, 0.793, label="Multi-atlas", size=4, hjust=0)+
-  coord_cartesian(ylim=c(0.7,0.8), xlim=c(3,18)) + 
-  scale_y_continuous(breaks=seq(0,1,0.01))
-
-@  
+\includegraphics{paper-001}
 \end{minipage}

### 2012-10-29 paper.Rnw
Adding Mallar's comments

@@
 to the multi-atlas approach of 1.9\%.
+{I STILL THINK THAT HIS SENTENCE NEEDS TO BE CLEARER; BREAK THIS OUT INTO TWO SENTENCES - OR MORE- WHERE FIRST YOU TALK ABOUT THE VALIDATION APPROACH, INCLUDING THE DIFFERENT METHODOLOGIES.
+IN A SECOND SENTENCE YOU SHOULD DISCUSS THE RESULTS.
+ALSO I THINK WE NEED SOME EMPHASIS HERE ON HOW MUCH BETTER THIS APPROACH IS THAN THE REGULAR NAIVE/ATLAS-BASED PROCEDURES.
+THERE IS A CLEAR BENEFIT TO USING MAGET BRAIN WHEN YOU DON'T WANT TO INVEST THE TIME INTO CREATING A TEMPLATE LIBRARY.}
 % fix last line: clarify "best case percent difference"
@@
 schizophrenia, and has functional significance in cognitive processes such as
-learning and memory.  For many research questions involving neuroimaging data,
+learning and memory.  For many research questions involving neuroimaging
+data, {CHANGE FROM MAGNETIC NEUROIMAGING TO MAGNETIC RESONANCE IMAGING
+(MRI) DATA}
 accurate identification of the hippocampus and its subregions in participant MR
 images is a necessary first step in order to analyse neuroanatomical
-differences and changes in sample populations.
+differences and changes in sample populations.  {CHANGE FROM "IN ORDER TO
+ANALYZE NEUROANAOMTICAL DIFFERENCES" TO "TO BETTER UNDERSTAND THE
+INDIVIDUAL NEUROANATOMY IN SPECIFIC SUBJECTS"}
 
@@
 development and aging process and across the time course of different diseases
-is expected. As well, it may be necessary to use several different hippocampal
+is expected.
+{CHANGE FROM "THE DEVELOPMENT AND AGING PROCESS AND ACROSS THE TIME COURSE OF DIFFERENT DISEASES IS EXPECTED" TO "DIFFERENTIAL EFFECTS  DEVELOPMENT AND AGING THROUGHOUT THE COURSE OF DIFFERENT NEUROPSYCHIATRIC DISORDERS ARE TO BE EXPECTED"--ALSO MAYBE INCLUDE A REFERENCE HERE}.
+As well, it may be necessary to use several different hippocampal
 definitions or, in fact, make research-specific modifications in order to test
@@
 Automated segmentation techniques attempt to address the need for human
-expertise by performing segmentations computationally.  A popular class of
+expertise by performing segmentations computationally.
+{NOT SURE THAT I TOTALLY UNDERSTAND THIS SENTENCE  -- MAYBE SOMETHING TO THE EFFECT OF "AUTOMATED SEGMENTATION TECHNIQUES WERE INITIALLY DESIGNED TO OVERCOME MANY OF THE LIMITATIONS INHERENT IN MANUAL SEGMENTATION PROCEDURES"} A popular class of
 automated methods, {\it atlas-based segmentation}, rely on a set of expertly
-labeled neuroanatomical atlases.  Each atlas is warped to fit a subject's
+labeled neuroanatomical atlases.
+{I'M GUESSING WE'RE JUMPING STRAIGHT TO MULTI-ATLAS HERE AND NOT ATLAS-BASED (GIVEN THE NEXT FEW LINES).
+IF THAT'S THE CASE WE MAY WANT TO ACTUALLY INCLUDE SOME BITS HERE ABOUT STRAIGHT UP ATLAS-BASED SEGMENTATION AND BRIEFLY DESCRIBE THEIR LIMITATIONS (HOWEVER WE MAYBE BOUNDED BY SPACE}  Each atlas is warped to fit a subject's
 neuroanatomy using nonlinear registration
@@
 
-Many descriptions of atlas-based segmentation algorithms report relying on an
+Many descriptions of atlas-based  {ATLAS OR MULTI-ATLAS??} segmentation algorithms report relying on an
 atlas library containing between 30 and 80 expertly labeled
@@
 In this paper we propose an automated segmentation technique to address the
-issues found in existing atlas-based methods. Principly, our method aims to
+issues found in existing atlas-based methods  {AND CREATE AND IMPROVEMENT OVER REGULAR ATLAS-BASED PROCEDURES??}.. Principly, our method aims to
 dramatically reduce the number of manually labelled atlases necessary.  It does
@@
 now-labelled image into the atlas library, but requires 30 starting atlases.
-Our contribution is provide a method that provides comparable segmentation
-accuracy to these and other atlas-based maethods whilst using significantly
+Our contribution is provide a method that provides {HOW ABOUT "OUTPUTS" RATHER THAN "PROVIDES"} comparable segmentation
+accuracy to these and other atlas-based  {ATLAS OR MULTI-ATLAS??} methods whilst using significantly
 fewer manually created atlases.
@@
 globus pallidus, and thalamus) using a single histologically-derived atlas.
+{ONE OF THE COMMENTS IN THE MICCAI REVIEWS WAS THAT THIS WAS A MODEST INCREMENTAL ADVANCE.
+YOU'VE DONE A TON OF VALIDATION HERE AND I WOULD DEFINITELY ARGUE THAT THIS FAR MORE THAN A MODEST ADVANCE.
+CAN WE ADD A LINE OR TWO HERE THAT BRINGS THE CONTRIBUTION HOME HERE??}
 In this work we extend our approach to the hippocampus, and validate it by
@@
 if it were the atlas library.  To do this we select unlabelled images from the
-input images to form the template library.  Next, labels from each atlas image
+input images to form the template library.
+{HAMMER HOME THAT THERE ARE DISTINCT ADVANTAGES TO THIS APPROACH- ie: USING THE UNIQUE POPULATION ON HAND TO BOOTSTRAP YOUR ANALYSIS}  Next, labels from each atlas image
 are propagated to each image in the template library (via the transformation
@@
 
-We use the {\it cross-correlation weighted voting} method to fuse the
-candidate labels of each unlabeled image.  In this method, a voxel-wise
-majority vote is carried out the labels from the top $n$ ranked template
-library images.  The scores are given by the normalized cross-correlation
-between the unlabeled image and each template library image, after the
-unlabeled image is linearly registered to the template image.  The
-cross-correlation score itself is calculated over a region of interest
-(ROI) defined on each template.  A customized ROI is defined by first
-linearly registering each unlabeled subject to all possible templates in
-the template library, transforming all template hippocampus labels,
-defining the possible extent of all hippocampal labels, and finally
-dilating this extent by three voxels.
-\todo{jp: it isn't clear that only the top ranked images are being used.
-perhaps some rationale for CC voting would help here. i.e. discuss improvement
-in overall registration quality --  throwing out atlases liable to introduce
-unwanted variance }
+We use the {\it cross-correlation weighted voting} method to both improve the
+quality of the label fusion step and reduce the number of registrations
+required \todo{ref}.  In this method, each template library image is ranked 
+in similarity to each unlabelled image, and only the top $n$ template library
+images labels are used for the voxel-wise majority vote fusion step.
+Similarity scores are computed by the normalized cross-correlation of image
+intensities between an unlabeled image and each template library image, after
+the unlabeled image has undergone linearly registration to the template image.
+The cross-correlation score is calculated over a region of interest (ROI) on
+each template representing an generous approximation of hippocampal structure
+and neighbourhood.  The ROI is defined as the extent of all atlas labels after
+linear registration to the template, dilated by three voxels. {IF THERE IS SPACE MENTION THAT THIS IS A SOMEWHAT AD-HOC HEURISTIC BUT AKIN TO WHAT OTHERS HAVE DONE LIKE HECKEMANN}
 
@@
 parameters optimized in Robbins et al.\cite{Robbins2004}.
+{PERHAPS ALSO MENTION THAT THE CHOICE OF ALGORITHM IS PRETTY WELL IRRELEVANT; AND YOU CAN CITE THE ORIGINAL MAGET BRAIN PAPER HERE - HERE WE SIMPLY CHOOSE AN ALGORITHM THAT WE KNOW BEST (OF COURSE SAY THIS A WEE BIT BETTER THAN THAT}
 
@@
 Massachusetts General Hospital and are available at
-\url{http://www.cma.mgh.harvard.edu/ibsr/}.
+\url{''http://www.cma.mgh.harvard.edu/ibsr/''}. {MAKE THE FACT THAT THERE ARE MANUAL SEGMENTATIONS ALREADY AVAILABLE OBVIOUS HERE}
 
@@
 
-In our experiments we verified how varying two different parameters effects
-labeling accuracy.
-
-\subsubsection{Varying Number of Atlases.}
-First we varied the number of input atlases that could be used from IBSR
-dataset from 3 to 8.  In each case, we conducted 30 rounds of random
-subsampling cross-validation using the 18 manual segmented templates from
-the IBSR dataset as input. In each round, atlases were randomly chosen from
-the IBSR dataset and the remaining templates are used both as template
+We explored how varying the size of the atlas library and the template library
+effects labeling accuracy.  For each parameter setting we conducted 30 rounds
+of random subsampling cross-validation using the 18 manual segmented templates
+from the IBSR dataset as input. In each round, atlases were randomly chosen
+from the IBSR dataset and the remaining images are used both as template
 library images and unlabeled subjects to be labeled using the MAGeT Brain
-algorithm.
-
-\subsubsection{Varying Number of Templates and Cross-Correlation Voting.}
-The second parameter we varied was the number of possible candidate
-labels to fuse for the final segmentation.  For each round of
-validation in Experiment 1, we carried out cross-correlation weighted
-label fusion amongst only the labels from the $n$ candidate templates
-from the library (i.e. selecting only those labels derived from
-templates with highest cross-correlation within the hippocampal ROI).
-The number of templates ($n$) was varied in the range $[3, r]$ (where $r$
-is the number of possible templates in the library).
-
-%where candidate labels are ranked by the cross-correlation score of the
-%subject to the template whence the labels came. We use values of n in the
-
-%In this experiment we compare the performance of MAGeT using three manually
-%segmented high-resolution atlases. ??We conduct a single run of the MAGeT
-%algorithm on 60 subjects and corresponding hippocampal segmentations from
-%the ADNI database. 20 subjects with AD, MCI and normals are used to
-%ftemplates from a pool of 60 templates to use in the template library, and
-%as target subjects. ??The pool of 60 templates is itself, randomly selected
-%from the ADNI database of first screening structural MRIs. ??
+algorithm.  We varied the size of the atlas library from 3 to 8, and used
+cross-correlation weighted label fusion to select the top $n$ candidate
+templates from the remaining images.  The size of the template library ($n$)
+was varied in the range $[3, 18-a]$ (where $a$ is the size of the atlas
+library).
 
@@
 Each segmentation was evaluated against the manual segmentation from the IBSR
-dataset using the Dice Kappa ($\kappa$) overlap metric:
-
-\[ \kappa = \frac{2a}{2a+b+c} \]
-
-\noindent where $a$ is the number of voxels common to the candidate
-segmentation and the gold standard and $b+c$ is the sum of the voxels
-uniquely identified by either the automatically generated candidate
-segmentation or the gold-standard.
+dataset using the Dice Kappa ($\kappa$) overlap metric, $\kappa =
+\frac{2a}{2a+b+c}$, where $a$ is the number of voxels common to the candidate
+segmentation and the gold standard and $b+c$ is the sum of the voxels uniquely
+identified by either the automatically generated candidate segmentation or the
+gold-standard.
 
@@
 reduction in false negatives (voxels labeled by the `gold-standard' only)
-in the hippocampal tail, a constant number of false positives (voxels
+in the hippocampal tail, a constant {HOW DO YOU KNOW IT'S CONSTANT ??}  number of false positives (voxels
 labeled by the MAGeT Brain method only) in the hippocampal head, and an
@@
 ($0.775$ for MAGeT Brain with 8 atlases and $0.790$ for multi-atlas after
-selecting the top 14 templates).
+selecting the top 14 templates).  {SOMETHING IN HERE AS AN EASY WAY TO BOOST ACCURACY FROM THE "ATLAS BASED CASE WOULD BE GOOD TOO}
 
@@
 mean naive segmentation, and mean multi-atlas with cross-correlation voting.}
-Multi-atlas and naive segmentations procedures provide upper and lower
-bounds (respectively) for accuracy.  Note that improved segmentation
-accuracy is achieved through increasing the number of input atlases.}
+Multi-atlas and naive segmentations procedures provide upper and lower bounds
+(respectively) for accuracy.  Note that improved segmentation accuracy is
+achieved through increasing the number of input atlases.}
 \label{results}

### 2012-10-29 paper.tex
Adding Mallar's comments

@@
 and neighbourhood.  The ROI is defined as the extent of all atlas labels after
-linear registration to the template, dilated by three voxels.
+linear registration to the template, dilated by three voxels. {IF THERE IS SPACE MENTION THAT THIS IS A SOMEWHAT AD-HOC HEURISTIC BUT AKIN TO WHAT OTHERS HAVE DONE LIKE HECKEMANN}
 
@@
 Massachusetts General Hospital and are available at
-\url{''http://www.cma.mgh.harvard.edu/ibsr/''}.
+\url{''http://www.cma.mgh.harvard.edu/ibsr/''}. {MAKE THE FACT THAT THERE ARE MANUAL SEGMENTATIONS ALREADY AVAILABLE OBVIOUS HERE}
 

### 2012-10-30 paper.Rnw
Incorporated most of Mallar's comments.

@@
 validate our approach and find an average best case percent difference
-to the multi-atlas approach of 1.9\%.
-{I STILL THINK THAT HIS SENTENCE NEEDS TO BE CLEARER; BREAK THIS OUT INTO TWO SENTENCES - OR MORE- WHERE FIRST YOU TALK ABOUT THE VALIDATION APPROACH, INCLUDING THE DIFFERENT METHODOLOGIES.
-IN A SECOND SENTENCE YOU SHOULD DISCUSS THE RESULTS.
-ALSO I THINK WE NEED SOME EMPHASIS HERE ON HOW MUCH BETTER THIS APPROACH IS THAN THE REGULAR NAIVE/ATLAS-BASED PROCEDURES.
-THERE IS A CLEAR BENEFIT TO USING MAGET BRAIN WHEN YOU DON'T WANT TO INVEST THE TIME INTO CREATING A TEMPLATE LIBRARY.}
+to the multi-atlas approach of 1.9\%. \todo{i still think that his sentence
+needs to be clearer; break this out into two sentences - or more- where
+first you talk about the validation approach, including the different
+methodologies.  in a second sentence you should discuss the results.  also
+i think we need some emphasis here on how much better this approach is than
+the regular naive/multi-atlas-based procedures.  there is a clear benefit to
+using maget brain when you don't want to invest the time into creating a
+template library.}
 % fix last line: clarify "best case percent difference"
@@
 implicated in forms of brain dysfunction such as Alzheimer's disease and
-schizophrenia, and has functional significance in cognitive processes such as
-learning and memory.  For many research questions involving neuroimaging
-data, {CHANGE FROM MAGNETIC NEUROIMAGING TO MAGNETIC RESONANCE IMAGING
-(MRI) DATA}
-accurate identification of the hippocampus and its subregions in participant MR
-images is a necessary first step in order to analyse neuroanatomical
-differences and changes in sample populations.  {CHANGE FROM "IN ORDER TO
-ANALYZE NEUROANAOMTICAL DIFFERENCES" TO "TO BETTER UNDERSTAND THE
-INDIVIDUAL NEUROANATOMY IN SPECIFIC SUBJECTS"}
+schizophrenia, and has functional significance in cognitive processes such
+as learning and memory.  For many research questions involving magnetic
+resonance imaging (MRI) data accurate identification of the hippocampus and
+its subregions in participant MR images is a necessary first step to better
+understand the individual neuroanatomy of specific subjects.
 
@@
 disputed\cite{Geuze2004}, as evidenced by efforts to create an unified
-segmentation protocol\cite{Jack2011}.  Compounding each of these problems is
-that significant neuroanatomical variability in the hippocampus throughout the
-development and aging process and across the time course of different diseases
-is expected.
-{CHANGE FROM "THE DEVELOPMENT AND AGING PROCESS AND ACROSS THE TIME COURSE OF DIFFERENT DISEASES IS EXPECTED" TO "DIFFERENTIAL EFFECTS  DEVELOPMENT AND AGING THROUGHOUT THE COURSE OF DIFFERENT NEUROPSYCHIATRIC DISORDERS ARE TO BE EXPECTED"--ALSO MAYBE INCLUDE A REFERENCE HERE}.
+segmentation protocol\cite{Jack2011}.  Compounding each of these problems
+is the significant neuroanatomical variability in the hippocampus
+throughout the course of aging, development, and neuropsychiatric
+disorders\cite{someone}.
+
 As well, it may be necessary to use several different hippocampal
@@
 
-Automated segmentation techniques attempt to address the need for human
-expertise by performing segmentations computationally.
-{NOT SURE THAT I TOTALLY UNDERSTAND THIS SENTENCE  -- MAYBE SOMETHING TO THE EFFECT OF "AUTOMATED SEGMENTATION TECHNIQUES WERE INITIALLY DESIGNED TO OVERCOME MANY OF THE LIMITATIONS INHERENT IN MANUAL SEGMENTATION PROCEDURES"} A popular class of
-automated methods, {\it atlas-based segmentation}, rely on a set of expertly
-labeled neuroanatomical atlases.
-{I'M GUESSING WE'RE JUMPING STRAIGHT TO MULTI-ATLAS HERE AND NOT ATLAS-BASED (GIVEN THE NEXT FEW LINES).
-IF THAT'S THE CASE WE MAY WANT TO ACTUALLY INCLUDE SOME BITS HERE ABOUT STRAIGHT UP ATLAS-BASED SEGMENTATION AND BRIEFLY DESCRIBE THEIR LIMITATIONS (HOWEVER WE MAYBE BOUNDED BY SPACE}  Each atlas is warped to fit a subject's
-neuroanatomy using nonlinear registration
-techniques\cite{Collins1995,Klein2009}.  Atlas labels are then transformed by
-this warping and a {\it label fusion} technique, such as voxel-wise voting, is
-used to merge the labellings from each atlas into a final segmentation for a
-subject.
+Automated segmentation techniques attempt to overcome the need for human
+expertise by performing segmentations computationally.  A popular class of
+automated methods, {\it multi-atlas-based segmentation}, rely on a set of
+expertly labeled neuroanatomical atlases. Each atlas is warped to fit a
+subject's neuroanatomy using nonlinear registration
+techniques\cite{Collins1995,Klein2009}.  Atlas labels are then transformed
+by this warping and a {\it label fusion} technique, such as voxel-wise
+voting, is used to merge the labellings from each atlas into a final
+segmentation for a subject.
 
-Many descriptions of atlas-based  {ATLAS OR MULTI-ATLAS??} segmentation algorithms report relying on an
+Many descriptions of multi-atlas-based segmentation algorithms report relying on an
 atlas library containing between 30 and 80 expertly labeled
@@
 In this paper we propose an automated segmentation technique to address the
-issues found in existing atlas-based methods  {AND CREATE AND IMPROVEMENT OVER REGULAR ATLAS-BASED PROCEDURES??}.. Principly, our method aims to
-dramatically reduce the number of manually labelled atlases necessary.  It does
-this by using a small atlas set to {\it generate} a much larger ``template
-library'', which is then used to segment the subjects in the same fashion as
-other atlas-based methods.  The essential insight of generating a template
-library is not new.  Heckemann \cite{Heckemann2006} compared generating a
-template library from a single atlas to standard multi-atlas segmentation and
-found poor performance and so deemed the approach as inviable.  The LEAP
-algorithm \cite{Wolz2010} proceeds by iteratively segmenting the unlabelled
-image most similar to the atlas library images and then incorporating the
-now-labelled image into the atlas library, but requires 30 starting atlases.
-Our contribution is provide a method that provides {HOW ABOUT "OUTPUTS" RATHER THAN "PROVIDES"} comparable segmentation
-accuracy to these and other atlas-based  {ATLAS OR MULTI-ATLAS??} methods whilst using significantly
-fewer manually created atlases.
+issues found in existing multi-atlas-based methods  \todo{AND CREATE AND
+IMPROVEMENT OVER REGULAR ATLAS-BASED PROCEDURES??}.. Principly, our method
+aims to dramatically reduce the number of manually labelled atlases
+necessary.  It does this by using a small atlas set to {\it generate} a
+much larger ``template library'', which is then used to segment the
+subjects in the same fashion as other multi-atlas-based methods.  This
+approach has the additional advantage of using the unique subject
+population on hand to bootstrap the segmentation analysis.
+
+The essential insight of generating a template library is not new.
+Heckemann \cite{Heckemann2006} compared generating a template library from
+a single atlas to standard multi-atlas segmentation and found poor
+performance and so deemed the approach as inviable.  The LEAP algorithm
+\cite{Wolz2010} proceeds by iteratively segmenting the unlabelled image
+most similar to the atlas library images and then incorporating the
+now-labelled image into the atlas library, but requires 30 starting
+atlases.  Our contribution is a method that outputs comparable segmentation
+accuracy to these and other multi-atlas-based methods whilst using
+significantly fewer manually created atlases.
 
@@
 globus pallidus, and thalamus) using a single histologically-derived atlas.
-{ONE OF THE COMMENTS IN THE MICCAI REVIEWS WAS THAT THIS WAS A MODEST INCREMENTAL ADVANCE.
-YOU'VE DONE A TON OF VALIDATION HERE AND I WOULD DEFINITELY ARGUE THAT THIS FAR MORE THAN A MODEST ADVANCE.
-CAN WE ADD A LINE OR TWO HERE THAT BRINGS THE CONTRIBUTION HOME HERE??}
-In this work we extend our approach to the hippocampus, and validate it by
-varying the atlas and template library sizes.  Due to the small number of
-atlases required, our method could easily accommodate different hippocampal
+\todo{challenge the "MODEST INCREMENTAL ADVANCE" argument} In this work we
+extend our approach to the hippocampus, and validate it by varying the
+atlas and template library sizes.  Due to the small number of atlases
+required, our method could easily accommodate different hippocampal
 definitions.  Our aim is not to improve on segmentation accuracy beyond
@@
 Our extension to this approach is to add a preliminary stage in which we
-construct a template library to be used in the standard atlas-based approach as
-if it were the atlas library.  To do this we select unlabelled images from the
-input images to form the template library.
-{HAMMER HOME THAT THERE ARE DISTINCT ADVANTAGES TO THIS APPROACH- ie: USING THE UNIQUE POPULATION ON HAND TO BOOTSTRAP YOUR ANALYSIS}  Next, labels from each atlas image
-are propagated to each image in the template library (via the transformation
-resulting from a non-linear registration between each pair of images).
-Thus, each template library image has a labelling from each atlas.  Traditional
-multi-atlas segmentation is then used to produce segmentations for the entire
-set of unlabelled images (including those images used in the template library).
+construct a template library to be used as if it were the atlas library
+in the standard multi-atlas-based method.  To do this we select from the
+input images to form the template library.  Next, labels from each atlas
+image are propagated to each image in the template library (via the
+transformation resulting from a non-linear registration between each pair
+of images).  Thus, each template library image has a labelling from each
+atlas.  Traditional multi-atlas segmentation is then used to produce
+segmentations for the entire set of unlabelled images (including those
+images used in the template library).
 
@@
 the unlabeled image has undergone linearly registration to the template image.
-The cross-correlation score is calculated over a region of interest (ROI) on
-each template representing an generous approximation of hippocampal structure
-and neighbourhood.  The ROI is defined as the extent of all atlas labels after
-linear registration to the template, dilated by three voxels. {IF THERE IS SPACE MENTION THAT THIS IS A SOMEWHAT AD-HOC HEURISTIC BUT AKIN TO WHAT OTHERS HAVE DONE LIKE HECKEMANN}
+The cross-correlation score is calculated over a region of interest (ROI)
+on each template, representing a generous approximation of hippocampal
+structure and neighbourhood, defined as the extent of all atlas labels
+after linear registration to the template, dilated by three voxels.
+\todo{ref Heckmann}.
 
@@
 coefficient. For the purposes of this work we used the regularization
-parameters optimized in Robbins et al.\cite{Robbins2004}.
-{PERHAPS ALSO MENTION THAT THE CHOICE OF ALGORITHM IS PRETTY WELL IRRELEVANT; AND YOU CAN CITE THE ORIGINAL MAGET BRAIN PAPER HERE - HERE WE SIMPLY CHOOSE AN ALGORITHM THAT WE KNOW BEST (OF COURSE SAY THIS A WEE BIT BETTER THAN THAT}
+parameters optimized in Robbins et al.\cite{Robbins2004}. It should be
+noted that the MAGeT brain algorithm is not dependant on this, or any,
+particular choice of registration method\cite{Chakravarty2011}.
 
@@
 Massachusetts General Hospital and are available at
-\url{''http://www.cma.mgh.harvard.edu/ibsr/''}. {MAKE THE FACT THAT THERE ARE MANUAL SEGMENTATIONS ALREADY AVAILABLE OBVIOUS HERE}
+\url{''http://www.cma.mgh.harvard.edu/ibsr/''}.
 
@@
 reduction in false negatives (voxels labeled by the `gold-standard' only)
-in the hippocampal tail, a constant {HOW DO YOU KNOW IT'S CONSTANT ??}  number of false positives (voxels
+in the hippocampal tail, a constant \todo{HOW DO YOU KNOW IT'S CONSTANT ??}  number of false positives (voxels
 labeled by the MAGeT Brain method only) in the hippocampal head, and an
@@
 
-% TODO: check nomenclature for cross-correlation label fusion (and other
-% methods) 
 \begin{figure}[h]
@@
 ($0.775$ for MAGeT Brain with 8 atlases and $0.790$ for multi-atlas after
-selecting the top 14 templates).  {SOMETHING IN HERE AS AN EASY WAY TO BOOST ACCURACY FROM THE "ATLAS BASED CASE WOULD BE GOOD TOO}
+selecting the top 14 templates).  \todo{SOMETHING IN HERE AS AN EASY WAY TO BOOST ACCURACY FROM THE "ATLAS BASED CASE WOULD BE GOOD TOO}
 

### 2012-10-31 paper.Rnw
Piles of edits incorporating more of Mallars comments and clarifications to the results and discussion section.

@@
 
-\newcommand{\todo}[1]{
-    \textcolor{red}{#1}
-}
+\newcommand{\todo}[1]{\textcolor{red}{#1}}
+\newcommand{\magetbrain}{MAGeT brain }
 
@@
 BRAIN ALGORITHM} 
-\name{  Jon Pipitone$^{\star}$ Jason P. Lerch$^{\dagger}$ Miriam Friedel$^{\dagger}$ Aristotle Voineskos$^{\star}$
-         Mallar Chakravarty$^{\star}$ }
+\name{  Jon Pipitone$^{1}$ Jason P. Lerch$^{4,5}$ Miriam Friedel$^{4,5}$
+        Aristotle Voineskos$^{1,3}$ Mallar Chakravarty$^{1,2,3}$ }
 
-\address{$^{\star}$ Centre for Additions and Mental Health, Toronto ON, Canada \\
-     $^{\dagger}$ Mouse Imaging Centre, ..., Toronto ON, Canada}
+\address{
+$^1$ Kimel Family Translational Imaging Genetics Research Laboratory, \\
+     Research Imaging Centre, Centre for Addiction and Mental Health, Toronto, ON, Canada \\
+$^2$ Institute of Biomaterials and Biomedical Engineering, University of
+     Toronto, Toronto, ON, Canada \\
+$^3$ Department of Psychiatry, University of Toronto, Toronto, ON, Canada \\
+$^4$ Program in Neuroscience and Mental Health, The Hospital for Sick Children,
+     Toronto, ON, Canada \\
+$^5$ Department of Medical Biophysics, University of Toronto, Toronto, ON, Canada
+}
 
@@
 \begin{abstract}
-Neuroimaging research often relies on automated anatomical
-segmentations of MR images of the brain. Current multi-atlased based
-approaches provide accurate segmentations of brain images by
-propagating region labels from manual delineations to unlabeled
-images.  Unfortunately, these approaches often rely on a large number
-of such manual segmentations which take time and significant expertise
-to produce, neither of which may be readily available.  We present
-an algorithm for the automatic segmentation of the hippocampus that
-minimizes the number of atlases needed whilst still achieving nearly
-the same accuracy as other multi-atlas approaches.  We perform
-repeated random sub-sampling validation on the IBSR dataset to
-validate our approach and find an average best case percent difference
-to the multi-atlas approach of 1.9\%. \todo{i still think that his sentence
-needs to be clearer; break this out into two sentences - or more- where
-first you talk about the validation approach, including the different
-methodologies.  in a second sentence you should discuss the results.  also
-i think we need some emphasis here on how much better this approach is than
-the regular naive/multi-atlas-based procedures.  there is a clear benefit to
-using maget brain when you don't want to invest the time into creating a
-template library.}
-% fix last line: clarify "best case percent difference"
+Neuroimaging research often relies on automated anatomical segmentations of MR
+images of the brain. Current multi-atlas based approaches provide accurate
+segmentations of brain images by propagating region labels from manual
+delineations to unlabeled images. Unfortunately, these approaches often rely on
+a large number of such manual-segmentated atlases which take time and
+significant expertise to produce. We present an algorithm for the automatic
+segmentation of the hippocampus that minimizes the number of atlases needed
+whilst still achieving similar accuracy to  other multi-atlas approaches.  We
+perform repeated random subsampling validation on the IBSR dataset to compare
+our approach to multi-atlas segmentation using the full IBSR dataset, and to
+single-atlas (model-based) segmentation. Our results show that with only 8
+input atlas, \magetbrain can achieve to within 2.0\% segmentation accuracy of
+the multi-atlas approach using 17 input atlases (mean $\kappa = 0.775$ vs.
+$\kappa = 0.791$).
 
@@
 implicated in forms of brain dysfunction such as Alzheimer's disease and
-schizophrenia, and has functional significance in cognitive processes such
-as learning and memory.  For many research questions involving magnetic
-resonance imaging (MRI) data accurate identification of the hippocampus and
-its subregions in participant MR images is a necessary first step to better
+schizophrenia, and has functional significance in cognitive processes such as
+learning and memory.  For many research questions involving magnetic resonance
+imaging (MRI) data accurate identification of the hippocampus and its
+subregions in participant MR images is a necessary first step to better
 understand the individual neuroanatomy of specific subjects.
@@
 In this paper we propose an automated segmentation technique to address the
-issues found in existing multi-atlas-based methods  \todo{AND CREATE AND
-IMPROVEMENT OVER REGULAR ATLAS-BASED PROCEDURES??}.. Principly, our method
-aims to dramatically reduce the number of manually labelled atlases
-necessary.  It does this by using a small atlas set to {\it generate} a
-much larger ``template library'', which is then used to segment the
-subjects in the same fashion as other multi-atlas-based methods.  This
-approach has the additional advantage of using the unique subject
-population on hand to bootstrap the segmentation analysis.
+above issues found in existing multi- atlas-based methods. Principly, our
+method aims to dramatically reduce the number of manually labelled atlases
+needed (under 10). It does this by using the small atlas library to boot-strap
+a much larger "template library", which is then used to segment the subjects in
+similar fashion as other multi-atlas-based methods. Our approach has the
+additional advantage of using the unique subject population on hand to
+initialise the segmentation analysis.
 
-The essential insight of generating a template library is not new.
-Heckemann \cite{Heckemann2006} compared generating a template library from
-a single atlas to standard multi-atlas segmentation and found poor
-performance and so deemed the approach as inviable.  The LEAP algorithm
-\cite{Wolz2010} proceeds by iteratively segmenting the unlabelled image
-most similar to the atlas library images and then incorporating the
-now-labelled image into the atlas library, but requires 30 starting
-atlases.  Our contribution is a method that outputs comparable segmentation
-accuracy to these and other multi-atlas-based methods whilst using
-significantly fewer manually created atlases.
+The essential insight of generating a template library is not new.  Heckemann
+\cite{Heckemann2006} compared generating a template library from a single atlas
+to standard multi-atlas segmentation and found poor performance and so deemed
+the approach as inviable.  The LEAP algorithm \cite{Wolz2010} proceeds by
+iteratively segmenting the unlabelled image most similar to the atlas library
+images and then incorporating the now-labelled image into the atlas library,
+but requires 30 starting atlases.  The novelty of our method is to demonstrate
+the possibility of producing comparable segmentation accuracy to these and
+other multi-atlas-based methods whilst using significantly fewer manually
+created atlases.
 
-In previous work from our group \cite{Chakravarty2011}, we explored
-this same approach to the segmentation of certain subcortical structures (striatum,
-globus pallidus, and thalamus) using a single histologically-derived atlas.
-\todo{challenge the "MODEST INCREMENTAL ADVANCE" argument} In this work we
-extend our approach to the hippocampus, and validate it by varying the
-atlas and template library sizes.  Due to the small number of atlases
-required, our method could easily accommodate different hippocampal
-definitions.  Our aim is not to improve on segmentation accuracy beyond
-existing methods, but instead to provide a method that trades off manual
+In previous work from our group \cite{Chakravarty2011}, we applied \magetbrain
+to the segmentation the striatum, globus pallidus, and thalamus using a single
+histologically-derived atlas. In the present work we extend our approach to the
+hippocampus, a much more difficult structure to segment, and perform a thorough
+validation over a range of atlas and template library sizes. Due to the small
+number of atlases required, our method could easily accommodate different
+hippocampal definitions. Our aim is not to improve on segmentation accuracy
+beyond existing methods, but instead to provide a method that trades off manual
 segmentation expertise for computational processing time whilst providing
@@
 
-We use the {\it cross-correlation weighted voting} method to both improve the
-quality of the label fusion step and reduce the number of registrations
-required \todo{ref}.  In this method, each template library image is ranked 
-in similarity to each unlabelled image, and only the top $n$ template library
-images labels are used for the voxel-wise majority vote fusion step.
-Similarity scores are computed by the normalized cross-correlation of image
-intensities between an unlabeled image and each template library image, after
-the unlabeled image has undergone linearly registration to the template image.
-The cross-correlation score is calculated over a region of interest (ROI)
-on each template, representing a generous approximation of hippocampal
-structure and neighbourhood, defined as the extent of all atlas labels
-after linear registration to the template, dilated by three voxels.
-\todo{ref Heckmann}.
+Label fusion is performed by cross-correlation weighted voting, a strategy
+weighted towards an optimal combination of subjects from the template library
+which has been previously shown to improve segmentation
+accuracy\cite{Aljabar2009,Collins2010}. In this method, each template library
+image is ranked in similarity to each unlabelled image by the normalized
+cross-correlation of image intensities after linear registration in a region of
+interest (ROI) generously encompassing hippocampus.  Only the top n ranked
+template library image labels are used in a voxel-wise majority vote. The ROI
+is heuristically defined as the extent of all atlas labels after linear
+registration to the template, dilated by three voxels\cite{Chakravarty2012}.
+
+\subsection{MRI dataset evaluated}
+
+For evaluation purposes we used the publicly available IBSR dataset.  This
+dataset consists of T1 weighted MR image volumes from 18 subjects (4 females,
+14 males) with ages between 7 and 71 years. Image dimensions for all MR volumes
+are normalized to  $256  \times 256 \times 128$ voxels, with the voxel size
+ranging from $0.8 \times 0.8 \times 1.5 mm$ to $1.0 \times 1.0 \times 1.5 mm$.
+The images come 'positionally normalized' into the Talairach orientation
+(rotation only), and processed by the CMA 'autoseg' biasfield correction
+routines. The MR brain data sets and their manual segmentations are publicly
+available and were provided by the Center for Morphometric Analysis at
+Massachusetts General Hospital and are available at
+\url{''http://www.cma.mgh.harvard.edu/ibsr/''}.
 
@@
 parameters optimized in Robbins et al.\cite{Robbins2004}. It should be
-noted that the MAGeT brain algorithm is not dependant on this, or any,
+noted that the \magetbrain algorithm is not dependant on this, or any,
 particular choice of registration method\cite{Chakravarty2011}.
-
-\subsection{MRI dataset evaluated}
-
-For evaluation purposes we used the publicly available IBSR dataset.  This
-dataset consists of T1 weighted MR image volumes from 18 subjects (4 females,
-14 males) with ages between 7 and 71 years. Image dimensions for all MR volumes
-are normalized to  $256  \times 256 \times 128$ voxels, with the voxel size
-ranging from $0.8 \times 0.8 \times 1.5 mm$ to $1.0 \times 1.0 \times 1.5 mm$.
-The images come 'positionally normalized' into the Talairach orientation
-(rotation only), and processed by the CMA 'autoseg' biasfield correction
-routines. The MR brain data sets and their manual segmentations are publicly
-available and were provided by the Center for Morphometric Analysis at
-Massachusetts General Hospital and are available at
-\url{''http://www.cma.mgh.harvard.edu/ibsr/''}.
 
@@
 from the IBSR dataset and the remaining images are used both as template
-library images and unlabeled subjects to be labeled using the MAGeT Brain
+library images and unlabeled subjects to be labeled using the \magetbrain
 algorithm.  We varied the size of the atlas library from 3 to 8, and used
@@
 The resulting segmentations from each of our experiments is compared to two
-alternative segmentation approaches. The {\it naive} approach is that in  which
-a single atlas is used to segment an unlabeled subject by directly propagating
-labels.  The {\it multi-atlas} approach, as described above, in which we used
-cross-correlation to select the top most similar atlases to each subject a
-atlas
-labels are propagated to an unlabeled subject and the labels are fused using
-the cross-correlation voting described above.  In total we evaluated
-approximately $52,000$ segmentations for the work presented in this manuscript.
+alternative segmentation approaches. The {\it single-atlas} approach uses one
+atlas to segment an unlabelled subject by directly propagating labels from
+atlas by way of nonlinear registration.  For each image in the IBSR dataset we
+computed the single-atlas segmentation from each of the other 17 labelled
+images. For the {\it multi-atlas} approach, described above, we performed
+leave-one-out validation in which each image in the IBSR dataset is segmented
+using an atlas library of the remaining 17 images. To compare with the MAGeT
+brain results we also varied the number of atlas images used in the label
+fusion step by employing cross-correlation weighted voting.  In total we
+evaluated approximately $52,000$ segmentations for the work presented in this
+manuscript.
 
@@
 
-Sample segmentations from one of the IBSR subjects is shown in Fig.
-\ref{montage}.  The segmentations shown are varied across the template
-selection for a case where 3 atlases were used.  The figure demonstrates a
-reduction in false negatives (voxels labeled by the `gold-standard' only)
-in the hippocampal tail, a constant \todo{HOW DO YOU KNOW IT'S CONSTANT ??}  number of false positives (voxels
-labeled by the MAGeT Brain method only) in the hippocampal head, and an
-increase in the segmentation accuracy as the number of templates used from
-the template library are increased.
+Sample segmentations from a single IBSR subject compared with the gold-standard
+segmentation are in Fig.  \ref{montage}.  As the size of the template library
+is increased, the number of false negatives in the hippocampal tail region is
+reduced.  This is correlated with an increase in segmentation accuracy.
 
@@
 \end{minipage}
-\caption{{\em Sample MAGeT Brain segmentations.} Segmentations of a single
-subject are shown when 3 atlases are used, with cross-correlation label
-fusion. Blue represents agreement between the gold-standard and MAGeT
-Brain.  Red indicates voxels labeled only by the gold-standard, and green
-indicates voxels labeled only by MAGeT Brain. Note improved accuracy as the
-number of templates is increased.} 
+\caption{{\em Sample \magetbrain segmentations.} 
+Segmentations of a single subject are shown when three atlases are used, with
+varying template library size $n$. Blue colouring represents agreement between
+the gold-standard and \magetbrain. Green colouring indicates false positive
+voxels labelled by \magetbrain, and red colouring indicates false negative
+voxels.
+} 
 \label{montage}
@@
 
-In Fig. \ref{results} we compare the performance of the MAGeT Brain method
-across all validations (i.e. a combination of all atlases and template
-selections) against the multi-atlas method and the naive approach.
-Regardless of the number of atlases and templates used in a particular
-segmentation we see a marked increase in segmentation accuracy over the
-average naive segmentation (the average Kappa of all segmentations derived
-from all pairwise registrations in the IBSR dataset).  Clearly, including
-more input atlases increases segmentation accuracy.  Surprisingly,
-increasing the number of labels in the label fusion step through
-cross-correlation based voting, does little to improve segmentation
-accuracy (except for in the cases where 3 and 4 atlases are used).
-Finally, the best-case MAGeT Brain performance scenario is when 8 atlases
-are used.  While this does not reach the accuracy of the multi-atlas
-segmentation, there is only a 1.9\% difference in the average Kappa values
-($0.775$ for MAGeT Brain with 8 atlases and $0.790$ for multi-atlas after
-selecting the top 14 templates).  \todo{SOMETHING IN HERE AS AN EASY WAY TO BOOST ACCURACY FROM THE "ATLAS BASED CASE WOULD BE GOOD TOO}
+\magetbrain shows distinct improvement in segmentation accuracy with increasing
+atlas library size (Fig. \ref{results}), reaching its highest accuracy using 8
+atlases (mean $\kappa = 0.775$).  This level of accuracy is within 2.0\% of the
+multi-atlas approach (mean $\kappa = 0.791$) using 17 atlases. With only 3
+atlases, \magetbrain reaches to within 6.5\% accuracy of the multi-atlas
+approach, and performs significantly better than the average single-atlas
+performance. Surprisingly, our results do not show any significant improvements
+in segmentation accuracy when applying cross-correlation weighted voting to
+reduce the number of template labels being fused.
 
@@
 qplot(x=top_n, y=k_hippocampus, data=xcorr_vote, colour=as.factor(num_atlases), geom=c('point'), alpha=I(0)) +
-  geom_smooth(line=1.5, formula=y~poly(x,2), method='lm') + 
+  geom_smooth(method="loess") + #line=1.5, formula=y~poly(x,2), method='lm') + 
   geom_hline(aes(yintercept = mean(subset(results, method=='naive')$k_hippocampus))) + 
   theme(legend.position="none") + 
-  ylab('Kappa') + 
-  xlab('Number of Atlases') +
-  annotate("text", 16.0, 0.712, label="Naive", size=4, hjust=0)+
+  ylab('Mean similarity score (Kappa)') + 
+  xlab('Number of template labels fused') +
+  annotate("text", 14.3, 0.712, label="Single-atlas", size=4, hjust=0)+
   annotate("text", 15.3, 0.754, label="3 atlases", size=4, hjust=0)+
@@
   annotate("text", 10.3, 0.779, label="8 atlases", size=4, hjust=0)+
-  annotate("text", 14.6, 0.793, label="Multi-atlas", size=4, hjust=0)+
-  coord_cartesian(ylim=c(0.7,0.8), xlim=c(3,18)) + 
+  annotate("text", 14.5, 0.793, label="Multi-atlas", size=4, hjust=0)+
+  annotate("text", 14.6, 0.787, label="17 atlases", size=4, hjust=0)+
+  scale_colour_manual(values=c(
+    "#D73027","#FC8D59","#FEE08B","#D9EF8B", "#91CF60","#1A9850", "#000000")) + 
+  coord_cartesian(ylim=c(0.703,0.803), xlim=c(3,18)) + 
   scale_y_continuous(breaks=seq(0,1,0.01))
@@
 \end{minipage}
-\caption{{\em Performance of MAGeT Brain across all validations compared to
-mean naive segmentation, and mean multi-atlas with cross-correlation voting.}
-Multi-atlas and naive segmentations procedures provide upper and lower bounds
-(respectively) for accuracy.  Note that improved segmentation accuracy is
-achieved through increasing the number of input atlases.}
+\caption{
+Mean performance of \magetbrain with varying atlas library size and number of
+template labels fused. Also shown is the mean multi-atlas performance when
+using an atlas library of 17 images and varying the number of labels fused, as
+well as the mean performance of single-atlas segmentations.  Data is fit with
+LOESS local regression smoothing lines, and include a standard deviation
+ribbon.
+}
 \label{results}
@@
 
-In this paper, we have demonstrated that accurate segmentations can be
-achieved by simply automatically deriving a template library from a small
-set of input atlases.  MAGeT Brain segmentations were compared to both
-naive segmentations and multi-atlas segmentations with cross-correlation
-voting.  In \cite{Lotjonen2010}, the authors report a Kappa of 0.814 on
-hippocampal segmentation in the IBSR dataset using a multi-atlas approach
-where atlas selection is based on the similarity between atlas and
-unlabeled subject after nonlinear registration, and using STAPLE
-\cite{Warfield2004} for label fusion.  We demonstrated that the trade off
-in accuracy over a multi-atlas segmentation using the entire IBSR manually
-segmented library is only on average 1.9\%, using only 8 randomly chosen
-atlases.  Discrepancies between our results and the above results may be
-due to choice of registration algorithm, regularization parameters, or
-similarity metric for label fusion.  Performance may also be affected by
-the variability in the IBSR data set (as previously noted by
-\cite{Klein2009}).  Future work from our group will attempt to address some
-of these issues.
+In this paper, we have demonstrated that accurate segmentations can be produced
+by automatically deriving a template library from a small set of input atlases.
+\magetbrain segmentations were compared to both single-atlas segmentations and
+multi-atlas segmentations with cross-correlation voting.  For the IBSR dataset,
+on average \magetbrain achieves within 2.0\% of the segmentation accuracy of
+the multi-atlas approach but requires only 8 atlases as compared to using the
+entire IBSR manually segmented library of 17 atlases.
 
-\todo{discuss CC voting and why accuracy is monotonically increasing...
-wouldn't we expect an increase and then decrease?}
+L\"{o}tj\"{o}nen et al.\cite{Lotjonen2010} report a Kappa of 0.814 on
+hippocampal segmentation in the IBSR dataset using a multi-atlas approach where
+atlas selection is based on the similarity between atlas and unlabeled subject
+after nonlinear registration, and using STAPLE\cite{Warfield2004} for label
+fusion.  Discrepancies between our results and the above results may be due to
+choice of registration algorithm, regularization parameters, or similarity
+metric for label fusion.  Performance may also be affected by the variability
+in the IBSR data set (as previously noted by \cite{Klein2009}).  Future work
+from our group will attempt to address some of these issues.
+
+\subsection{Acknowledgements}
+Computations were performed on the gpc supercomputer at the SciNet HPC
+Consortium\cite{Loken2010}. SciNet is funded by: the Canada Foundation for
+Innovation under the auspices of Compute Canada; the Government of Ontario;
+Ontario Research Fund - Research Excellence; and the University of Toronto.
 

### 2012-10-31 paper.Rnw
Fixes and concision

@@
 \newcommand{\todo}[1]{\textcolor{red}{#1}}
-\newcommand{\magetbrain}{MAGeT brain }
 
@@
 single-atlas (model-based) segmentation. Our results show that with only 8
-input atlas, \magetbrain can achieve to within 2.0\% segmentation accuracy of
+input atlas, MAGeT brain can achieve to within 2.0\% segmentation accuracy of
 the multi-atlas approach using 17 input atlases (mean $\kappa = 0.775$ vs.
@@
 
-In previous work from our group \cite{Chakravarty2011}, we applied \magetbrain
+In previous work from our group \cite{Chakravarty2011}, we applied MAGeT brain
 to the segmentation the striatum, globus pallidus, and thalamus using a single
@@
 The segmentation approach we propose is best understood as an extension of
-traditional multi-atlas segmentation.  In multi-atlas segmentation, the inputs
-are an atlas library and a set of unlabelled MR images.  An unlabelled image is
-segmented in the following way.  Each atlas image is non-linearly registered to
-the unlabelled image, and then each atlas' labels are propagated via the
-resulting transformation.  These candidate labels are then fused to produce a
-definitive segmentation. The label fusion method used may determine which
-templates are used to segment a particular subject (for instance, voxel-wise
-majority vote takes labels from all atlases, whereas a weighted majority vote
-selects only those templates most similar to the subject).
+traditional multi-atlas segmentation.  In multi-atlas segmentation, an atlas
+library and unlabelled MR images are given as input.  Every atlas image is
+nonlinearly registered to each unlabelled image, and then each atlas' labels
+are propagated via the resulting transformations.  These labels are then fused
+to produce a single, definitive segmentation by some label fusion method (e.g.
+voxel-wise majority vote).
 
-Our extension to this approach is to add a preliminary stage in which we
-construct a template library to be used as if it were the atlas library
-in the standard multi-atlas-based method.  To do this we select from the
-input images to form the template library.  Next, labels from each atlas
-image are propagated to each image in the template library (via the
-transformation resulting from a non-linear registration between each pair
-of images).  Thus, each template library image has a labelling from each
-atlas.  Traditional multi-atlas segmentation is then used to produce
-segmentations for the entire set of unlabelled images (including those
-images used in the template library).
+Our extension adds a preliminary stage in which a template library is
+constructed from input images, and used in place of an atlas library in the
+standard multi-atlas-based method.  To create the template library, labels from
+each atlas image are propagated to each template library image via the
+transformation resulting from a non-linear registration between pair of images.
+As a result, each template library image has a label from each atlas.
+Traditional multi-atlas segmentation is then used to produce segmentations for
+the entire set of unlabelled images (including those images used in the
+template library).
 
@@
 Massachusetts General Hospital and are available at
-\url{''http://www.cma.mgh.harvard.edu/ibsr/''}.
+\href{http://www.cma.mgh.harvard.edu/ibsr/}
+{\nolinkurl{http://www.cma.mgh.harvard.edu/ibsr/}}.
 
@@
 The N3 algorithm \cite{Sled1998} is first used to minimize the intensity
-nonuniformity in each of the atlases and unlabeled subject images.  Image
-registration is carried out in two phases.  In the first, a 12-parameter linear
-transformation (3 translations, rotations, scales, shears) is estimated between
-images using an algorithm that maximizes the correlation between blurred MR
-intensities and gradient magnitude over the whole brain \cite{Collins}.  In the
-second phase, nonlinear registration is completed using the ANIMAL algorithm
-\cite{Collins1995}: an iterative procedure that estimates a 3D deformation
-field between two MR images. At first, large deformations are estimated using
-blurred version of the input data. These larger deformations are then input to
-subsequent steps where the fit is refined by estimating smaller deformations on
-data blurred with a Gaussian kernel with a smaller FWHM. The final
-transformation is a set of local translations defined on a bed of equally
-spaced nodes that were estimated through the optimization of the correlation
-coefficient. For the purposes of this work we used the regularization
-parameters optimized in Robbins et al.\cite{Robbins2004}. It should be
-noted that the \magetbrain algorithm is not dependant on this, or any,
-particular choice of registration method\cite{Chakravarty2011}.
+nonuniformity in each image.  Image registration is carried out in two phases.
+In the first, a 12-parameter linear transformation (3 translations, rotations,
+scales, shears) is estimated between images using an algorithm that maximizes
+the correlation between blurred MR intensities and gradient magnitude over the
+whole brain \cite{Collins}.  In the second phase, nonlinear registration is
+completed using the ANIMAL algorithm \cite{Collins1995}: an iterative procedure
+that estimates a 3D deformation field between two MR images. At first, large
+deformations are estimated using blurred version of the input data. These
+larger deformations are then input to subsequent steps where the fit is refined
+by estimating smaller deformations on data blurred with a Gaussian kernel with
+a smaller FWHM. The final transformation is a set of local translations defined
+on a bed of equally spaced nodes that were estimated through the optimization
+of the correlation coefficient. For the purposes of this work we used the
+regularization parameters optimized in Robbins et al.\cite{Robbins2004}. It
+should be noted that the MAGeT brain algorithm is not dependant on this, or
+any, particular choice of registration method\cite{Chakravarty2011}.
 
@@
 from the IBSR dataset and the remaining images are used both as template
-library images and unlabeled subjects to be labeled using the \magetbrain
+library images and unlabeled subjects to be labeled using the MAGeT brain
 algorithm.  We varied the size of the atlas library from 3 to 8, and used
 cross-correlation weighted label fusion to select the top $n$ candidate
-templates from the remaining images.  The size of the template library ($n$)
-was varied in the range $[3, 18-a]$ (where $a$ is the size of the atlas
-library).
+templates from the remaining images.  $n$ was varied in the range $[3, 18-a]$,
+where $a$ is the size of the atlas library.
 
@@
 \subsubsection{Goodness-of-fit}
-Each segmentation was evaluated against the manual segmentation from the IBSR
-dataset using the Dice Kappa ($\kappa$) overlap metric, $\kappa =
-\frac{2a}{2a+b+c}$, where $a$ is the number of voxels common to the candidate
-segmentation and the gold standard and $b+c$ is the sum of the voxels uniquely
-identified by either the automatically generated candidate segmentation or the
-gold-standard.
+Automatically produced segmentations are evaluated against IBSR manual
+segmentations dataset using the Dice Kappa ($\kappa$) overlap metric, $\kappa =
+{2a}/{(2a+b+c)}$, where $a$ is the number of voxels common to both
+segmentations and $b+c$ is the sum of the voxels uniquely identified in either
+segmentation.
 
 \subsubsection{Comparison Approaches}
-The resulting segmentations from each of our experiments is compared to two
-alternative segmentation approaches. The {\it single-atlas} approach uses one
-atlas to segment an unlabelled subject by directly propagating labels from
-atlas by way of nonlinear registration.  For each image in the IBSR dataset we
-computed the single-atlas segmentation from each of the other 17 labelled
-images. For the {\it multi-atlas} approach, described above, we performed
-leave-one-out validation in which each image in the IBSR dataset is segmented
-using an atlas library of the remaining 17 images. To compare with the MAGeT
-brain results we also varied the number of atlas images used in the label
-fusion step by employing cross-correlation weighted voting.  In total we
-evaluated approximately $52,000$ segmentations for the work presented in this
-manuscript.
+The resulting segmentations from each of our experiments are compared to those
+produced from two alternative segmentation approaches. The {\it single-atlas}
+approach uses one atlas to segment a unlabelled subject by directly propagating
+labels from the atlas by way of nonlinear registration.  We computed a
+single-atlas segmentation for each image in the IBSR dataset from each of the
+other 17 labelled images.  Similarly, we computed a segmentation for each image
+using the {\it multi-atlas} approach, described above, using the other 17
+images as the atlas library.  Additionally, we also varied the number of atlas
+images used in the label fusion step by employing cross-correlation weighted
+voting.  In total we evaluated approximately $52,000$ segmentations for the
+work presented in this manuscript.
 
@@
 \end{minipage}
-\caption{{\em Sample \magetbrain segmentations.} 
+\caption{{\em Sample MAGeT brain segmentations.} 
 Segmentations of a single subject are shown when three atlases are used, with
 varying template library size $n$. Blue colouring represents agreement between
-the gold-standard and \magetbrain. Green colouring indicates false positive
-voxels labelled by \magetbrain, and red colouring indicates false negative
+the gold-standard and MAGeT brain. Green colouring indicates false positive
+voxels labelled by MAGeT brain, and red colouring indicates false negative
 voxels.
@@
 
-\magetbrain shows distinct improvement in segmentation accuracy with increasing
+MAGeT brain shows distinct improvement in segmentation accuracy with increasing
 atlas library size (Fig. \ref{results}), reaching its highest accuracy using 8
@@
 multi-atlas approach (mean $\kappa = 0.791$) using 17 atlases. With only 3
-atlases, \magetbrain reaches to within 6.5\% accuracy of the multi-atlas
+atlases, MAGeT brain reaches to within 6.5\% accuracy of the multi-atlas
 approach, and performs significantly better than the average single-atlas
@@
 \caption{
-Mean performance of \magetbrain with varying atlas library size and number of
+Mean performance of MAGeT brain with varying atlas library size and number of
 template labels fused. Also shown is the mean multi-atlas performance when
@@
 by automatically deriving a template library from a small set of input atlases.
-\magetbrain segmentations were compared to both single-atlas segmentations and
+MAGeT brain segmentations were compared to both single-atlas segmentations and
 multi-atlas segmentations with cross-correlation voting.  For the IBSR dataset,
-on average \magetbrain achieves within 2.0\% of the segmentation accuracy of
+on average MAGeT brain achieves within 2.0\% of the segmentation accuracy of
 the multi-atlas approach but requires only 8 atlases as compared to using the

### 2012-10-31 paper.tex
Fixes and concision

@@
-\documentclass{article}
+\documentclass{article}\usepackage{graphicx, color}
+%% maxwidth is the original width if it is less than linewidth
+%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
+\makeatletter
+\def\maxwidth{ %
+  \ifdim\Gin@nat@width>\linewidth
+    \linewidth
+  \else
+    \Gin@nat@width
+  \fi
+}
+\makeatother
+
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+\definecolor{fgcolor}{rgb}{0.2, 0.2, 0.2}
+\newcommand{\hlnumber}[1]{\textcolor[rgb]{0,0,0}{#1}}%
+\newcommand{\hlfunctioncall}[1]{\textcolor[rgb]{0.501960784313725,0,0.329411764705882}{\textbf{#1}}}%
+\newcommand{\hlstring}[1]{\textcolor[rgb]{0.6,0.6,1}{#1}}%
+\newcommand{\hlkeyword}[1]{\textcolor[rgb]{0,0,0}{\textbf{#1}}}%
+\newcommand{\hlargument}[1]{\textcolor[rgb]{0.690196078431373,0.250980392156863,0.0196078431372549}{#1}}%
+\newcommand{\hlcomment}[1]{\textcolor[rgb]{0.180392156862745,0.6,0.341176470588235}{#1}}%
+\newcommand{\hlroxygencomment}[1]{\textcolor[rgb]{0.43921568627451,0.47843137254902,0.701960784313725}{#1}}%
+\newcommand{\hlformalargs}[1]{\textcolor[rgb]{0.690196078431373,0.250980392156863,0.0196078431372549}{#1}}%
+\newcommand{\hleqformalargs}[1]{\textcolor[rgb]{0.690196078431373,0.250980392156863,0.0196078431372549}{#1}}%
+\newcommand{\hlassignement}[1]{\textcolor[rgb]{0,0,0}{\textbf{#1}}}%
+\newcommand{\hlpackage}[1]{\textcolor[rgb]{0.588235294117647,0.709803921568627,0.145098039215686}{#1}}%
+\newcommand{\hlslot}[1]{\textit{#1}}%
+\newcommand{\hlsymbol}[1]{\textcolor[rgb]{0,0,0}{#1}}%
+\newcommand{\hlprompt}[1]{\textcolor[rgb]{0.2,0.2,0.2}{#1}}%
+
+\usepackage{framed}
+\makeatletter
+\newenvironment{kframe}{%
+ \def\at@end@of@kframe{}%
+ \ifinner\ifhmode%
+  \def\at@end@of@kframe{\end{minipage}}%
+  \begin{minipage}{\columnwidth}%
+ \fi\fi%
+ \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
+ \colorbox{shadecolor}{##1}\hskip-\fboxsep
+     % There is no \\@totalrightmargin, so:
+     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
+ \MakeFramed {\advance\hsize-\width
+   \@totalleftmargin\z@ \linewidth\hsize
+   \@setminipage}}%
+ {\par\unskip\endMakeFramed%
+ \at@end@of@kframe}
+\makeatother
+
+\definecolor{shadecolor}{rgb}{.97, .97, .97}
+\definecolor{messagecolor}{rgb}{0, 0, 0}
+\definecolor{warningcolor}{rgb}{1, 0, 1}
+\definecolor{errorcolor}{rgb}{1, 0, 0}
+\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX
+
+\usepackage{alltt}
 \usepackage{spconf,amsmath,graphicx,color,hyperref}
@@
 
-\usepackage{Sweave}
 \begin{document}
-\input{paper-concordance}
+
 \maketitle
@@
   %\includegraphics[width=\textwidth]{results}
-\includegraphics{paper-001}
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\includegraphics[width=\maxwidth]{figure/unnamed-chunk-1} 
+\end{knitrout}
+
 \end{minipage}
 \caption{
-Mean performance of \magetbrain with varying atlas library size and number of
+Mean performance of MAGeT brain with varying atlas library size and number of
 template labels fused. Also shown is the mean multi-atlas performance when

### 2012-11-01 paper.Rnw
Incorporated Mallar's changes, and fixed dangling references.

@@
 
-\title{HIPPOCAMPAL SEGMENTATION WITH A VERY SMALL ATLAS LIBRARY: THE MAGET
+\title{HIPPOCAMPAL SEGMENTATION WITH A VERY SMALL ATLAS LIBRARY: THE MAGeT
 BRAIN ALGORITHM} 
-\name{  Jon Pipitone$^{1}$ Jason P. Lerch$^{4,5}$ Miriam Friedel$^{4,5}$
-        Aristotle Voineskos$^{1,3}$ Mallar Chakravarty$^{1,2,3}$ }
+\name{  Jon Pipitone$^{1}$ Jason P. Lerch$^{4,5}$ Miriam Friedel$^{4}$
+        Aristotle N. Voineskos$^{1,3}$ M. Mallar Chakravarty$^{1,2,3}$ }
 
@@
 images of the brain. Current multi-atlas based approaches provide accurate
-segmentations of brain images by propagating region labels from manual
-delineations to unlabeled images. Unfortunately, these approaches often rely on
-a large number of such manual-segmentated atlases which take time and
-significant expertise to produce. We present an algorithm for the automatic
-segmentation of the hippocampus that minimizes the number of atlases needed
-whilst still achieving similar accuracy to  other multi-atlas approaches.  We
-perform repeated random subsampling validation on the IBSR dataset to compare
-our approach to multi-atlas segmentation using the full IBSR dataset, and to
-single-atlas (model-based) segmentation. Our results show that with only 8
-input atlas, MAGeT brain can achieve to within 2.0\% segmentation accuracy of
-the multi-atlas approach using 17 input atlases (mean $\kappa = 0.775$ vs.
-$\kappa = 0.791$).
-
+segmentations of brain images by propagating manually derived segmentations of
+specific neuroanatomical structures to unlabelled data. Unfortunately, these
+approaches often rely on a large number of such manually segmentated atlases
+that take significant time and expertise to produce. We present an algorithm
+for the automatic segmentation of the hippocampus that minimizes the number of
+atlases needed whilst still achieving similar accuracy to multi-atlas
+approaches.  We perform repeated random subsampling validation on the IBSR
+dataset to compare our approach to multi-atlas segmentation using the full IBSR
+dataset, and to single-atlas (model-based) segmentation. Our results show that
+with only 8 input atlas, MAGeT brain can achieve to within 2.0\% segmentation
+accuracy of the multi-atlas approach using 17 input atlases (mean $\kappa =
+0.775$ vs.  $\kappa = 0.791$).  These results compare favourably to others who
+have used the IBSR data for validation. 
 \end{abstract}
@@
 identification of specific neuroanatomical structures by an expert manual
-rater.  Third, the true delination of hippocampal anatomy in MR images is
+rater.  Third, the true delineation of hippocampal anatomy in MR images is
 disputed\cite{Geuze2004}, as evidenced by efforts to create an unified
-segmentation protocol\cite{Jack2011}.  Compounding each of these problems
-is the significant neuroanatomical variability in the hippocampus
-throughout the course of aging, development, and neuropsychiatric
-disorders\cite{someone}.
+segmentation protocol\cite{Jack2011}.  Compounding each of these problems is
+the significant neuroanatomical variability in the hippocampus throughout the
+course of aging, development, and neuropsychiatric disorders\cite{Mouiha2011}.
 
@@
 
-In previous work from our group \cite{Chakravarty2011}, we applied MAGeT brain
+In previous work from our group \cite{Chakravarty2012}, we applied MAGeT brain
 to the segmentation the striatum, globus pallidus, and thalamus using a single
@@
 should be noted that the MAGeT brain algorithm is not dependant on this, or
-any, particular choice of registration method\cite{Chakravarty2011}.
+any, particular choice of registration method\cite{Chakravarty2012}.
 

### 2012-11-01 paper.Rnw
Miriam and Julie comments incorporated. Changed all discussion of multi-atlas to "basic multi-atlas" to disambiguate. Squashed into four pages.

@@
 for the automatic segmentation of the hippocampus that minimizes the number of
-atlases needed whilst still achieving similar accuracy to multi-atlas
+atlases needed while still achieving similar accuracy to multi-atlas
 approaches.  We perform repeated random subsampling validation on the IBSR
-dataset to compare our approach to multi-atlas segmentation using the full IBSR
-dataset, and to single-atlas (model-based) segmentation. Our results show that
-with only 8 input atlas, MAGeT brain can achieve to within 2.0\% segmentation
-accuracy of the multi-atlas approach using 17 input atlases (mean $\kappa =
-0.775$ vs.  $\kappa = 0.791$).  These results compare favourably to others who
-have used the IBSR data for validation. 
+dataset to compare our approach to basic multi-atlas segmentation using
+the full IBSR dataset, and to single-atlas (model-based) segmentation. Our
+results show that with only 8 input atlas, MAGeT brain can achieve to within
+2.0\% segmentation accuracy of the basic multi-atlas approach using 17
+input atlases (mean $\kappa = 0.775$ vs.  $\kappa = 0.791$).  These results
+compare favourably to others who have used the IBSR data for validation. 
 \end{abstract}
@@
 disputed\cite{Geuze2004}, as evidenced by efforts to create an unified
-segmentation protocol\cite{Jack2011}.  Compounding each of these problems is
-the significant neuroanatomical variability in the hippocampus throughout the
-course of aging, development, and neuropsychiatric disorders\cite{Mouiha2011}.
+segmentation protocol\cite{Jack2011}.
 
-As well, it may be necessary to use several different hippocampal
-definitions or, in fact, make research-specific modifications in order to test
-hypotheses (for example, \cite{Poppenk2011} found that overall hippocampal
-volume difference did not predict recollection memory performance but by
-dividing the hippocampus into anterior and posterior regions, a predictive
-volume difference was found).  Thus, whilst manual segmentation of the
-hippocampus is an important technique, it may be a bottleneck for researchers
-or clinicians who do not have access to the needed human expertise.
+Compounding each of these problems is the significant neuroanatomical
+variability in the hippocampus throughout the course of aging, development, and
+neuropsychiatric disorders\cite{Mouiha2011}.  Additionally, it may be necessary
+to use several different hippocampal definitions or, in fact, make
+research-specific modifications in order to test hypotheses (for example,
+\cite{Poppenk2011} found that overall hippocampal volume difference did not
+predict recollection memory performance but by dividing the hippocampus into
+anterior and posterior regions, a predictive volume difference was found).
+Thus, while manual segmentation of the hippocampus is an important technique,
+it may be a bottleneck for researchers or clinicians who do not have access to
+the needed human expertise.
 
@@
 the possibility of producing comparable segmentation accuracy to these and
-other multi-atlas-based methods whilst using significantly fewer manually
+other multi-atlas-based methods while using significantly fewer manually
 created atlases.
@@
 beyond existing methods, but instead to provide a method that trades off manual
-segmentation expertise for computational processing time whilst providing
+segmentation expertise for computational processing time while providing
 sufficient accuracy for clinical and research applications.
@@
 The segmentation approach we propose is best understood as an extension of
-traditional multi-atlas segmentation.  In multi-atlas segmentation, an atlas
+basic multi-atlas segmentation.  In multi-atlas segmentation, an atlas
 library and unlabelled MR images are given as input.  Every atlas image is
@@
 As a result, each template library image has a label from each atlas.
-Traditional multi-atlas segmentation is then used to produce segmentations for
-the entire set of unlabelled images (including those images used in the
-template library).
+Basic multi-atlas segmentation is then used to produce segmentations for the
+entire set of unlabelled images (including those images used in the template
+library).
 
@@
 other 17 labelled images.  Similarly, we computed a segmentation for each image
-using the {\it multi-atlas} approach, described above, using the other 17
-images as the atlas library.  Additionally, we also varied the number of atlas
-images used in the label fusion step by employing cross-correlation weighted
-voting.  In total we evaluated approximately $52,000$ segmentations for the
-work presented in this manuscript.
+using the {\it basic multi-atlas} approach, described above, using the
+other 17 images as the atlas library.  Additionally, we also varied the number
+of atlas images used in the label fusion step by employing cross-correlation
+weighted voting.  In total we evaluated approximately $52,000$ segmentations
+for the work presented in this manuscript.
 
@@
 \end{minipage}
-\caption{{\em Sample MAGeT brain segmentations.} 
+\caption{{\small
 Segmentations of a single subject are shown when three atlases are used, with
@@
 the gold-standard and MAGeT brain. Green colouring indicates false positive
-voxels labelled by MAGeT brain, and red colouring indicates false negative
-voxels.
-} 
+voxels labelled by MAGeT brain (i.e. labelled voxels not appearing in the
+gold-standard labels), and red colouring indicates false negative voxels (i.e.
+voxels labelled in the gold-standard but not by MAGeT brain).
+}} 
 \label{montage}
@@
 
-MAGeT brain shows distinct improvement in segmentation accuracy with increasing
-atlas library size (Fig. \ref{results}), reaching its highest accuracy using 8
-atlases (mean $\kappa = 0.775$).  This level of accuracy is within 2.0\% of the
-multi-atlas approach (mean $\kappa = 0.791$) using 17 atlases. With only 3
-atlases, MAGeT brain reaches to within 6.5\% accuracy of the multi-atlas
-approach, and performs significantly better than the average single-atlas
-performance. Surprisingly, our results do not show any significant improvements
-in segmentation accuracy when applying cross-correlation weighted voting to
-reduce the number of template labels being fused.
+MAGeT brain achieves a level of segmentation accuracy that is to within 2.0\%
+of the accuracy of the basic multi-atlas approach (Fig. \ref{results};
+mean $\kappa = 0.775$ vs mean $\kappa = 0.791$) in the best case. Importantly,
+to do this MAGeT brain only requires 8 manual segmented images whereas the
+basic multi-atlas requires 17 atlases.  This represents a significant
+savings in manual effort, and supports our contention that it is possible with
+MAGeT brain to trade a small decrease in accuracy for a significant decrease in
+the number of manual segmentations needed.
 
-\begin{figure}
-\begin{minipage}[b]{1.0\linewidth}
+Surprisingly, our analysis does not show any significant improvements in
+segmentation accuracy for either method when applying cross-correlation
+weighted voting to reduce the number of template labels being fused.
+
+\begin{figure}[t]
+%\begin{minipage}[b]{1.0\linewidth}
   \centering
   %\includegraphics[width=\textwidth]{results}
-<<fig=TRUE, cached=TRUE, echo=FALSE, warnings=FALSE, >>=
+<<fig=TRUE, cached=TRUE, echo=FALSE, warnings=FALSE >>=
 library(ggplot2)
@@
   annotate("text", 10.3, 0.779, label="8 atlases", size=4, hjust=0)+
-  annotate("text", 14.5, 0.793, label="Multi-atlas", size=4, hjust=0)+
+  annotate("text", 13.5, 0.793, label="Basic multi-atlas", size=4, hjust=0)+
   annotate("text", 14.6, 0.787, label="17 atlases", size=4, hjust=0)+
@@
 @  
-\end{minipage}
-\caption{
+%\end{minipage}
+\caption{{\small
 Mean performance of MAGeT brain with varying atlas library size and number of
-template labels fused. Also shown is the mean multi-atlas performance when
-using an atlas library of 17 images and varying the number of labels fused, as
-well as the mean performance of single-atlas segmentations.  Data is fit with
-LOESS local regression smoothing lines, and include a standard deviation
-ribbon.
-}
+template labels fused. Also shown is the mean basic multi-atlas
+performance when using an atlas library of 17 images and varying the number of
+labels fused, as well as the mean performance of single-atlas segmentations.
+Data is fit with LOESS local regression smoothing. One standard deviation is
+shown in grey.
+}}
 \label{results}
@@
 
-\subsection{Acknowledgements}
+\section{Acknowledgements}
 Computations were performed on the gpc supercomputer at the SciNet HPC

### 2012-11-01 paper.tex
Miriam and Julie comments incorporated. Changed all discussion of multi-atlas to "basic multi-atlas" to disambiguate. Squashed into four pages.

@@
 
-MAGeT brain shows distinct improvement in segmentation accuracy with increasing
-atlas library size (Fig. \ref{results}), reaching its highest accuracy using 8
-atlases (mean $\kappa = 0.775$).  This level of accuracy is within 2.0\% of the
-multi-atlas approach (mean $\kappa = 0.791$) using 17 atlases. With only 3
-atlases, MAGeT brain reaches to within 6.5\% accuracy of the multi-atlas
-approach, and performs significantly better than the average single-atlas
-performance. Surprisingly, our results do not show any significant improvements
-in segmentation accuracy when applying cross-correlation weighted voting to
-reduce the number of template labels being fused.
+MAGeT brain achieves a level of segmentation accuracy that is to within 2.0\%
+of the accuracy of the basic multi-atlas approach (Fig. \ref{results};
+mean $\kappa = 0.775$ vs mean $\kappa = 0.791$) in the best case. Importantly,
+to do this MAGeT brain only requires 8 manual segmented images whereas the
+basic multi-atlas requires 17 atlases.  This represents a significant
+savings in manual effort, and supports our contention that it is possible with
+MAGeT brain to trade a small decrease in accuracy for a significant decrease in
+the number of manual segmentations needed.
 
-\begin{figure}
-\begin{minipage}[b]{1.0\linewidth}
+Surprisingly, our analysis does not show any significant improvements in
+segmentation accuracy for either method when applying cross-correlation
+weighted voting to reduce the number of template labels being fused.
+
+\begin{figure}[t]
+%\begin{minipage}[b]{1.0\linewidth}
   \centering
@@
 
-\end{minipage}
-\caption{
+%\end{minipage}
+\caption{{\small
 Mean performance of MAGeT brain with varying atlas library size and number of
-template labels fused. Also shown is the mean multi-atlas performance when
-using an atlas library of 17 images and varying the number of labels fused, as
-well as the mean performance of single-atlas segmentations.  Data is fit with
-LOESS local regression smoothing lines, and include a standard deviation
-ribbon.
-}
+template labels fused. Also shown is the mean basic multi-atlas
+performance when using an atlas library of 17 images and varying the number of
+labels fused, as well as the mean performance of single-atlas segmentations.
+Data is fit with LOESS local regression smoothing. One standard deviation is
+shown in grey.
+}}
 \label{results}

### 2012-11-02 paper.Rnw
Merged in another round of mallar's changes.

@@
 imaging (MRI) data accurate identification of the hippocampus and its
-subregions in participant MR images is a necessary first step to better
-understand the individual neuroanatomy of specific subjects.
+subregions is a necessary first step to better understand the individual
+neuroanatomy of subjects.
 
@@
 by this warping and a {\it label fusion} technique, such as voxel-wise
-voting, is used to merge the labellings from each atlas into a final
-segmentation for a subject.
+voting, is used to merge the competing label definitions from each atlas into a
+final segmentation for a subject.
 
@@
 The segmentation approach we propose is best understood as an extension of
-basic multi-atlas segmentation.  In multi-atlas segmentation, an atlas
-library and unlabelled MR images are given as input.  Every atlas image is
-nonlinearly registered to each unlabelled image, and then each atlas' labels
-are propagated via the resulting transformations.  These labels are then fused
-to produce a single, definitive segmentation by some label fusion method (e.g.
-voxel-wise majority vote).
+basic multi-atlas segmentation\cite{Collins2010}.  In multi-atlas segmentation,
+an atlas library and unlabelled MR images are given as input.  Every atlas
+image is nonlinearly registered to each unlabelled image, and then each atlas'
+labels are propagated via the resulting transformations.  These labels are then
+fused to produce a single, definitive segmentation by some label fusion method
+(e.g.  voxel-wise majority vote).
 

### 2012-11-08 paper.Rnw
Last minute touch ups.  This is the version submitted to ISBI 1st round.
Also included here are the data from running an multi-atlas segmentation
with 9 atlases only.

@@
 \documentclass{article}
-\usepackage{spconf,amsmath,graphicx,color,hyperref}
+\usepackage{spconf,amsmath,graphicx,color}
+\usepackage[square,numbers]{natbib}
+\setlength{\bibsep}{0.0pt}
 
-\newcommand{\todo}[1]{\textcolor{red}{#1}}
-
-\title{HIPPOCAMPAL SEGMENTATION WITH A VERY SMALL ATLAS LIBRARY: THE MAGeT
-BRAIN ALGORITHM} 
+\title{How many atlases do you really need for accurate multi-atlas segmentation?} 
 \name{  Jon Pipitone$^{1}$ Jason P. Lerch$^{4,5}$ Miriam Friedel$^{4}$
@@
 segmentations of brain images by propagating manually derived segmentations of
-specific neuroanatomical structures to unlabelled data. Unfortunately, these
-approaches often rely on a large number of such manually segmentated atlases
-that take significant time and expertise to produce. We present an algorithm
-for the automatic segmentation of the hippocampus that minimizes the number of
-atlases needed while still achieving similar accuracy to multi-atlas
-approaches.  We perform repeated random subsampling validation on the IBSR
-dataset to compare our approach to basic multi-atlas segmentation using
-the full IBSR dataset, and to single-atlas (model-based) segmentation. Our
-results show that with only 8 input atlas, MAGeT brain can achieve to within
-2.0\% segmentation accuracy of the basic multi-atlas approach using 17
-input atlases (mean $\kappa = 0.775$ vs.  $\kappa = 0.791$).  These results
-compare favourably to others who have used the IBSR data for validation. 
+specific neuroanatomical structures to unlabelled data. These approaches often
+rely on a large number of such manually segmented atlases that take significant
+time and expertise to produce. We present an algorithm for the automatic
+segmentation of the hippocampus that minimizes the number of atlases needed
+while still achieving similar accuracy to multi-atlas approaches.  We perform
+repeated random subsampling validation on the International Brain Segmentation
+Repository (IBSR) dataset to compare our approach to basic multi-atlas
+segmentation using the full IBSR dataset, and to single-atlas (model-based)
+segmentation. Our results show that with only 8 input atlases, MAGeT brain can
+achieve to within 2.0\% segmentation accuracy of the basic multi-atlas approach
+using 17 input atlases (mean $\kappa = 0.775$ vs.  $\kappa = 0.791$).  These
+results compare favourably to those of other investigators who have used the IBSR
+data for validation.
 \end{abstract}
@@
 \begin{keywords}
-hippocampus, segmentation, multi-atlas, automated
+hippocampus, segmentation, multi-atlas methods, label fusion, magnetic
+resonance imaging 
 \end{keywords}
@@
 Currently, the gold standard for neuroanatomical segmentation is manual
-labelling by an expert human rater.  This is problematic for segmentation of
+delineation by an expert human rater.  This is problematic for segmentation of
 the hippocampus for several reasons.  First, manual segmentation takes a
-significant investment of time and expertise \cite{Hammers2003} which may not
+significant investment of time and expertise \citep{Hammers2003} which may not
 be readily available to researchers or clinicians.  Second, the amount of data
@@
 identification of specific neuroanatomical structures by an expert manual
-rater.  Third, the true delineation of hippocampal anatomy in MR images is
-disputed\cite{Geuze2004}, as evidenced by efforts to create an unified
-segmentation protocol\cite{Jack2011}.
+rater.  Third, the true definition of hippocampal anatomy in MR images is
+disputed \citep{Geuze2004}, as evidenced by efforts to create an unified
+segmentation protocol \citep{Jack2011}.
 
@@
 variability in the hippocampus throughout the course of aging, development, and
-neuropsychiatric disorders\cite{Mouiha2011}.  Additionally, it may be necessary
-to use several different hippocampal definitions or, in fact, make
-research-specific modifications in order to test hypotheses (for example,
-\cite{Poppenk2011} found that overall hippocampal volume difference did not
-predict recollection memory performance but by dividing the hippocampus into
-anterior and posterior regions, a predictive volume difference was found).
-Thus, while manual segmentation of the hippocampus is an important technique,
-it may be a bottleneck for researchers or clinicians who do not have access to
-the needed human expertise.
+neuropsychiatric disorders \citep{Mouiha2011}.  Additionally, it may be
+necessary to use several different hippocampal definitions or, in fact, make
+specific modifications in the course of research. For example, Poppenk et al.
+\citep{Poppenk2011} found that subdividing the hippocampus into anterior and
+posterior regions resulted in a predictive relationship between volume
+difference of those regions and recollection memory performance.  Thus, while
+manual segmentation of the hippocampus is a necessary technique, to researchers
+or clinicians who do not have access to the needed human expertise its use may
+be infeasible.
 
-Automated segmentation techniques attempt to overcome the need for human
-expertise by performing segmentations computationally.  A popular class of
-automated methods, {\it multi-atlas-based segmentation}, rely on a set of
-expertly labeled neuroanatomical atlases. Each atlas is warped to fit a
-subject's neuroanatomy using nonlinear registration
-techniques\cite{Collins1995,Klein2009}.  Atlas labels are then transformed
-by this warping and a {\it label fusion} technique, such as voxel-wise
-voting, is used to merge the competing label definitions from each atlas into a
-final segmentation for a subject.
+Automated segmentation techniques overcome the need for human expertise by
+performing segmentations computationally.  A popular class of automated
+methods, {\it multi-atlas-based segmentation}, rely on a set of expertly
+labeled neuroanatomical atlases. Each atlas is warped to fit a subject's
+neuroanatomy using nonlinear registration
+techniques \citep{Collins1995,Klein2009}.  Atlas labels are then transformed by
+this warping and a {\it label fusion} technique, such as voxel-wise voting, is
+used to merge the competing label definitions from each atlas into a final
+segmentation for a subject.
 
-Many descriptions of multi-atlas-based segmentation algorithms report relying on an
-atlas library containing between 30 and 80 expertly labeled
-brains\cite{Heckemann2011,Collins2010,Aljabar2009,Leung2010,Lotjonen2010}.
-As noted, the production of an atlas library requires significant manual
-effort, and is limited since the choice of atlases or segmentation protocol may
-not reflect the underlying neuroanatomical variability of the population under
+Many descriptions of multi-atlas-based segmentation algorithms report relying
+on an atlas library containing anywhere between 30 and 80 expertly labeled
+brains \citep{Heckemann2011,Collins2010,Aljabar2009,Leung2010,Lotjonen2010}.  As
+noted, the production of an atlas library requires significant manual effort,
+and is limited since the choice of atlases or segmentation protocol may not
+reflect the underlying neuroanatomical variability of the population under
 study or be suited to answer the research questions at hand.
 
-In this paper we propose an automated segmentation technique to address the
-above issues found in existing multi- atlas-based methods. Principly, our
-method aims to dramatically reduce the number of manually labelled atlases
-needed (under 10). It does this by using the small atlas library to boot-strap
-a much larger "template library", which is then used to segment the subjects in
-similar fashion as other multi-atlas-based methods. Our approach has the
-additional advantage of using the unique subject population on hand to
-initialise the segmentation analysis.
+In this paper we propose an automated segmentation method to address the above
+issues of existing multi-atlas-based methods. Principally, our method aims to
+dramatically reduce the number of manually labelled atlases needed (under 10).
+This is achieved by using the small atlas library to boot-strap a much larger
+"template library", which is then used to segment the subjects in a similar
+fashion to basic multi-atlas segmentation. This approach has the additional
+advantage of using the unique subject population on hand to initialize the
+segmentation process and improve accuracy.
 
 The essential insight of generating a template library is not new.  Heckemann
-\cite{Heckemann2006} compared generating a template library from a single atlas
+\citep{Heckemann2006} compared generating a template library from a single atlas
 to standard multi-atlas segmentation and found poor performance and so deemed
-the approach as inviable.  The LEAP algorithm \cite{Wolz2010} proceeds by
+the approach as inviable.  The LEAP algorithm \citep{Wolz2010} proceeds by
 iteratively segmenting the unlabelled image most similar to the atlas library
@@
 
-In previous work from our group \cite{Chakravarty2012}, we applied MAGeT brain
-to the segmentation the striatum, globus pallidus, and thalamus using a single
-histologically-derived atlas. In the present work we extend our approach to the
-hippocampus, a much more difficult structure to segment, and perform a thorough
-validation over a range of atlas and template library sizes. Due to the small
-number of atlases required, our method could easily accommodate different
-hippocampal definitions. Our aim is not to improve on segmentation accuracy
-beyond existing methods, but instead to provide a method that trades off manual
-segmentation expertise for computational processing time while providing
-sufficient accuracy for clinical and research applications.
+In our previous work \citep{Chakravarty2012}, we applied MAGeT brain to
+segmentation of the human striatum, globus pallidus, and thalamus using a
+single histologically-derived atlas. The main contribution of this paper is to
+extend our approach to the human hippocampus and perform a thorough validation
+over a range of atlas and template library sizes, which was not done in our
+previous work.  Due to the small number of atlases required, our method can
+easily accommodate different hippocampal definitions. Our aim is not to improve
+on segmentation accuracy beyond existing methods, but instead to provide a
+method that trades off manual segmentation expertise for computational
+processing time while providing sufficient accuracy for clinical and research
+applications.
 
@@
 The segmentation approach we propose is best understood as an extension of
-basic multi-atlas segmentation\cite{Collins2010}.  In multi-atlas segmentation,
+basic multi-atlas segmentation \citep{Collins2010}.  In multi-atlas segmentation,
 an atlas library and unlabelled MR images are given as input.  Every atlas
@@
 which has been previously shown to improve segmentation
-accuracy\cite{Aljabar2009,Collins2010}. In this method, each template library
+accuracy \citep{Aljabar2009,Collins2010}. In this method, each template library
 image is ranked in similarity to each unlabelled image by the normalized
 cross-correlation of image intensities after linear registration in a region of
-interest (ROI) generously encompassing hippocampus.  Only the top n ranked
+interest (ROI) generously encompassing the hippocampus.  Only the top ranked
 template library image labels are used in a voxel-wise majority vote. The ROI
 is heuristically defined as the extent of all atlas labels after linear
-registration to the template, dilated by three voxels\cite{Chakravarty2012}.
+registration to the template, dilated by three voxels \citep{Chakravarty2012}.
 
@@
 For evaluation purposes we used the publicly available IBSR dataset.  This
-dataset consists of T1 weighted MR image volumes from 18 subjects (4 females,
+dataset consists of T1-weighted MR image volumes from 18 subjects (4 females,
 14 males) with ages between 7 and 71 years. Image dimensions for all MR volumes
@@
 Massachusetts General Hospital and are available at
-\href{http://www.cma.mgh.harvard.edu/ibsr/}
-{\nolinkurl{http://www.cma.mgh.harvard.edu/ibsr/}}.
+{\tt http://www.cma.mgh.harvard.edu/ibsr/}.
 
 \subsection{Image Processing and Registration Method}
-The N3 algorithm \cite{Sled1998} is first used to minimize the intensity
+The N3 algorithm \citep{Sled1998} is first used to minimize the intensity
 nonuniformity in each image.  Image registration is carried out in two phases.
@@
 the correlation between blurred MR intensities and gradient magnitude over the
-whole brain \cite{Collins}.  In the second phase, nonlinear registration is
-completed using the ANIMAL algorithm \cite{Collins1995}: an iterative procedure
+whole brain \citep{Collins}.  In the second phase, nonlinear registration is
+completed using the ANIMAL algorithm \citep{Collins1995}: an iterative procedure
 that estimates a 3D deformation field between two MR images. At first, large
@@
 of the correlation coefficient. For the purposes of this work we used the
-regularization parameters optimized in Robbins et al.\cite{Robbins2004}. It
-should be noted that the MAGeT brain algorithm is not dependant on this, or
-any, particular choice of registration method\cite{Chakravarty2012}.
+regularization parameters optimized in Robbins et al. \citep{Robbins2004}. It
+should be noted that the MAGeT brain algorithm is not dependent on this, or
+any, particular choice of registration method \citep{Chakravarty2012}.
 
@@
 Sample segmentations from a single IBSR subject compared with the gold-standard
-segmentation are in Fig.  \ref{montage}.  As the size of the template library
-is increased, the number of false negatives in the hippocampal tail region is
-reduced.  This is correlated with an increase in segmentation accuracy.
+segmentation are in Fig.  \ref{montage}.  Qualitatively, as the size of the
+template library is increased, the number of false negatives in the hippocampal
+tail region is reduced, and segmentation accuracy increases.
 
@@
   %\includegraphics[width=\textwidth]{results}
-<<fig=TRUE, cached=TRUE, echo=FALSE, warnings=FALSE >>=
+<<performance, fig=TRUE, cached=TRUE, echo=FALSE, warnings=FALSE >>=
 library(ggplot2)
 results = read.csv(file='data/results.csv')
+#results = read.csv(file='data/results_with_9multiatlas.csv')
 xcorr_vote= subset(results, method=='xcorr' | method == 'multiatlas_xcorr_vote' )
 qplot(x=top_n, y=k_hippocampus, data=xcorr_vote, colour=as.factor(num_atlases), geom=c('point'), alpha=I(0)) +
-  geom_smooth(method="loess") + #line=1.5, formula=y~poly(x,2), method='lm') + 
-  geom_hline(aes(yintercept = mean(subset(results, method=='naive')$k_hippocampus))) + 
+  geom_smooth(method='loess')+ #line=1.5, formula=y~poly(x,2), method='lm') + 
+  geom_hline(aes(yintercept = mean(subset(results, method=='naive')$k_hippocampus), linetype="dashed")) + 
   theme(legend.position="none") + 
@@
     "#D73027","#FC8D59","#FEE08B","#D9EF8B", "#91CF60","#1A9850", "#000000")) + 
-  coord_cartesian(ylim=c(0.703,0.803), xlim=c(3,18)) + 
-  scale_y_continuous(breaks=seq(0,1,0.01))
-
+  scale_linetype_manual(values=c(
+    "dotted", "solid", "solid", "solid",
+         "solid", "solid", "solid", "dashed")) +
+  coord_cartesian(ylim=c(0.703,0.803), xlim=c(3,18)) + scale_y_continuous(breaks=seq(0,1,0.01))   
 @
@@
 
-L\"{o}tj\"{o}nen et al.\cite{Lotjonen2010} report a Kappa of 0.814 on
+L\"{o}tj\"{o}nen et al. \citep{Lotjonen2010} report a Kappa of 0.814 on
 hippocampal segmentation in the IBSR dataset using a multi-atlas approach where
 atlas selection is based on the similarity between atlas and unlabeled subject
-after nonlinear registration, and using STAPLE\cite{Warfield2004} for label
+after nonlinear registration, and using STAPLE \citep{Warfield2004} for label
 fusion.  Discrepancies between our results and the above results may be due to
@@
 metric for label fusion.  Performance may also be affected by the variability
-in the IBSR data set (as previously noted by \cite{Klein2009}).  Future work
+in the IBSR data set (as previously noted by \citep{Klein2009}).  Future work
 from our group will attempt to address some of these issues.
@@
 Computations were performed on the gpc supercomputer at the SciNet HPC
-Consortium\cite{Loken2010}. SciNet is funded by: the Canada Foundation for
+Consortium \citep{Loken2010}. SciNet is funded by: the Canada Foundation for
 Innovation under the auspices of Compute Canada; the Government of Ontario;
@@
 
+This work was supported by the Canadian Institutes of Health Research (CIHR),
+National Alliance for Research on Schizophrenia and Depression (NARSAD),
+Ontario Mental Health Foundation (OMHF), and the CAMH Foundation (Koerner New
+Scientist Program and Paul Garfinkel New Investigator Catalyst Fund).
 \bibliographystyle{IEEEbib}

### 2012-11-08 paper.tex
Last minute touch ups.  This is the version submitted to ISBI 1st round.
Also included here are the data from running an multi-atlas segmentation
with 9 atlases only.

@@
 \usepackage{alltt}
-\usepackage{spconf,amsmath,graphicx,color,hyperref}
+\usepackage{spconf,amsmath,graphicx,color}
+\usepackage[square,numbers]{natbib}
+\setlength{\bibsep}{0.0pt}
 
-\newcommand{\todo}[1]{\textcolor{red}{#1}}
-
-\title{HIPPOCAMPAL SEGMENTATION WITH A VERY SMALL ATLAS LIBRARY: THE MAGeT
-BRAIN ALGORITHM} 
+\title{How many atlases do you really need for accurate multi-atlas segmentation?} 
 \name{  Jon Pipitone$^{1}$ Jason P. Lerch$^{4,5}$ Miriam Friedel$^{4}$
@@
 \begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\includegraphics[width=\maxwidth]{figure/unnamed-chunk-1} 
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\includegraphics[width=\maxwidth]{figure/performance} 
 \end{knitrout}

### 2016-04-30 paper.Rnw
Updated to build with knitr 1.0

@@
 \begin{document}
-\SweaveOpts{concordance=TRUE}
+
+<<include=FALSE>>=
+library(knitr)
+opts_chunk$set(
+concordance=TRUE
+)
+@
+
 \maketitle
@@
   %\includegraphics[width=\textwidth]{results}
-<<performance, fig=TRUE, cached=TRUE, echo=FALSE, warnings=FALSE >>=
+<<performance, cached=TRUE, echo=FALSE, warnings=FALSE >>=
 library(ggplot2)

### 2016-04-30 paper.Rmd
First hack at a conversion to markdown

--- 2012-11-08 paper.tex
@@
-\documentclass{article}\usepackage{graphicx, color}
-%% maxwidth is the original width if it is less than linewidth
-%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
-\makeatletter
-\def\maxwidth{ %
-  \ifdim\Gin@nat@width>\linewidth
-    \linewidth
-  \else
-    \Gin@nat@width
-  \fi
-}
-\makeatother
+---
+title: "How many atlases do you really need for accurate multi-atlas segmentation?"
+author: Jon Pipitone, Jason P. Lerch, Miriam Friedel, Aristotle N. Voineskos, M. Mallar Chakravarty
+output: 
+    pdf_document:
+        fig_caption: true
+bibliography: references.bib
+---
 
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-\definecolor{fgcolor}{rgb}{0.2, 0.2, 0.2}
-\newcommand{\hlnumber}[1]{\textcolor[rgb]{0,0,0}{#1}}%
-\newcommand{\hlfunctioncall}[1]{\textcolor[rgb]{0.501960784313725,0,0.329411764705882}{\textbf{#1}}}%
-\newcommand{\hlstring}[1]{\textcolor[rgb]{0.6,0.6,1}{#1}}%
-\newcommand{\hlkeyword}[1]{\textcolor[rgb]{0,0,0}{\textbf{#1}}}%
-\newcommand{\hlargument}[1]{\textcolor[rgb]{0.690196078431373,0.250980392156863,0.0196078431372549}{#1}}%
-\newcommand{\hlcomment}[1]{\textcolor[rgb]{0.180392156862745,0.6,0.341176470588235}{#1}}%
-\newcommand{\hlroxygencomment}[1]{\textcolor[rgb]{0.43921568627451,0.47843137254902,0.701960784313725}{#1}}%
-\newcommand{\hlformalargs}[1]{\textcolor[rgb]{0.690196078431373,0.250980392156863,0.0196078431372549}{#1}}%
-\newcommand{\hleqformalargs}[1]{\textcolor[rgb]{0.690196078431373,0.250980392156863,0.0196078431372549}{#1}}%
-\newcommand{\hlassignement}[1]{\textcolor[rgb]{0,0,0}{\textbf{#1}}}%
-\newcommand{\hlpackage}[1]{\textcolor[rgb]{0.588235294117647,0.709803921568627,0.145098039215686}{#1}}%
-\newcommand{\hlslot}[1]{\textit{#1}}%
-\newcommand{\hlsymbol}[1]{\textcolor[rgb]{0,0,0}{#1}}%
-\newcommand{\hlprompt}[1]{\textcolor[rgb]{0.2,0.2,0.2}{#1}}%
+[^1]: Kimel Family Translational Imaging Genetics Research Laboratory, Research Imaging Centre, Centre for Addiction and Mental Health, Toronto, ON, Canada 
+[^2]: Institute of Biomaterials and Biomedical Engineering, University of Toronto, Toronto, ON, Canada
+[^3]: Department of Psychiatry, University of Toronto, Toronto, ON, Canada
+[^4]: Program in Neuroscience and Mental Health, The Hospital for Sick Children, Toronto, ON, Canada
+[^5]: Department of Medical Biophysics, University of Toronto, Toronto, ON, Canada
 
-\usepackage{framed}
-\makeatletter
-\newenvironment{kframe}{%
- \def\at@end@of@kframe{}%
- \ifinner\ifhmode%
-  \def\at@end@of@kframe{\end{minipage}}%
-  \begin{minipage}{\columnwidth}%
- \fi\fi%
- \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
- \colorbox{shadecolor}{##1}\hskip-\fboxsep
-     % There is no \\@totalrightmargin, so:
-     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
- \MakeFramed {\advance\hsize-\width
-   \@totalleftmargin\z@ \linewidth\hsize
-   \@setminipage}}%
- {\par\unskip\endMakeFramed%
- \at@end@of@kframe}
-\makeatother
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(cache=TRUE)
+```
 
-\definecolor{shadecolor}{rgb}{.97, .97, .97}
-\definecolor{messagecolor}{rgb}{0, 0, 0}
-\definecolor{warningcolor}{rgb}{1, 0, 1}
-\definecolor{errorcolor}{rgb}{1, 0, 0}
-\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX
-
-\usepackage{alltt}
-\usepackage{spconf,amsmath,graphicx,color}
-\usepackage[square,numbers]{natbib}
-\setlength{\bibsep}{0.0pt}
-
-\title{How many atlases do you really need for accurate multi-atlas segmentation?} 
-\name{  Jon Pipitone$^{1}$ Jason P. Lerch$^{4,5}$ Miriam Friedel$^{4}$
-        Aristotle N. Voineskos$^{1,3}$ M. Mallar Chakravarty$^{1,2,3}$ }
-
-\address{
-$^1$ Kimel Family Translational Imaging Genetics Research Laboratory, \\
-     Research Imaging Centre, Centre for Addiction and Mental Health, Toronto, ON, Canada \\
-$^2$ Institute of Biomaterials and Biomedical Engineering, University of
-     Toronto, Toronto, ON, Canada \\
-$^3$ Department of Psychiatry, University of Toronto, Toronto, ON, Canada \\
-$^4$ Program in Neuroscience and Mental Health, The Hospital for Sick Children,
-     Toronto, ON, Canada \\
-$^5$ Department of Medical Biophysics, University of Toronto, Toronto, ON, Canada
-}
-
-\begin{document}
-
-\maketitle
-
-\begin{abstract}
+# Abstract
 Neuroimaging research often relies on automated anatomical segmentations of MR
@@
 data for validation.
-\end{abstract}
 
-\begin{keywords}
-hippocampus, segmentation, multi-atlas methods, label fusion, magnetic
-resonance imaging 
-\end{keywords}
+**keywords:** hippocampus, segmentation, multi-atlas methods, label fusion,
+magnetic resonance imaging
 
-\section{Introduction}
-\label{sec:intro}
+Introduction {#sec:intro}
+============
 
-The hippocampus is of particular interest to many researchers because it is
-implicated in forms of brain dysfunction such as Alzheimer's disease and
-schizophrenia, and has functional significance in cognitive processes such as
-learning and memory.  For many research questions involving magnetic resonance
-imaging (MRI) data accurate identification of the hippocampus and its
-subregions is a necessary first step to better understand the individual
-neuroanatomy of subjects.
+The hippocampus is of particular interest to many researchers because it
+is implicated in forms of brain dysfunction such as Alzheimer’s disease
+and schizophrenia, and has functional significance in cognitive
+processes such as learning and memory. For many research questions
+involving magnetic resonance imaging (MRI) data accurate identification
+of the hippocampus and its subregions is a necessary first step to
+better understand the individual neuroanatomy of subjects.
 
 Currently, the gold standard for neuroanatomical segmentation is manual
-delineation by an expert human rater.  This is problematic for segmentation of
-the hippocampus for several reasons.  First, manual segmentation takes a
-significant investment of time and expertise \citep{Hammers2003} which may not
-be readily available to researchers or clinicians.  Second, the amount of data
-produced in neuroimaging experiments increasingly exceeds the capacity for
-identification of specific neuroanatomical structures by an expert manual
-rater.  Third, the true definition of hippocampal anatomy in MR images is
-disputed \citep{Geuze2004}, as evidenced by efforts to create an unified
-segmentation protocol \citep{Jack2011}.
+delineation by an expert human rater. This is problematic for
+segmentation of the hippocampus for several reasons. First, manual
+segmentation takes a significant investment of time and expertise
+[@Hammers2003] which may not be readily available to researchers or
+clinicians. Second, the amount of data produced in neuroimaging
+experiments increasingly exceeds the capacity for identification of
+specific neuroanatomical structures by an expert manual rater. Third,
+the true definition of hippocampal anatomy in MR images is disputed
+[@Geuze2004], as evidenced by efforts to create an unified segmentation
+protocol [@Jack2011].
 
 Compounding each of these problems is the significant neuroanatomical
-variability in the hippocampus throughout the course of aging, development, and
-neuropsychiatric disorders \citep{Mouiha2011}.  Additionally, it may be
-necessary to use several different hippocampal definitions or, in fact, make
-specific modifications in the course of research. For example, Poppenk et al.
-\citep{Poppenk2011} found that subdividing the hippocampus into anterior and
-posterior regions resulted in a predictive relationship between volume
-difference of those regions and recollection memory performance.  Thus, while
-manual segmentation of the hippocampus is a necessary technique, to researchers
-or clinicians who do not have access to the needed human expertise its use may
-be infeasible.
+variability in the hippocampus throughout the course of aging,
+development, and neuropsychiatric disorders [@Mouiha2011]. Additionally,
+it may be necessary to use several different hippocampal definitions or,
+in fact, make specific modifications in the course of research. For
+example, Poppenk et al. [@Poppenk2011] found that subdividing the
+hippocampus into anterior and posterior regions resulted in a predictive
+relationship between volume difference of those regions and recollection
+memory performance. Thus, while manual segmentation of the hippocampus
+is a necessary technique, to researchers or clinicians who do not have
+access to the needed human expertise its use may be infeasible.
 
-Automated segmentation techniques overcome the need for human expertise by
-performing segmentations computationally.  A popular class of automated
-methods, {\it multi-atlas-based segmentation}, rely on a set of expertly
-labeled neuroanatomical atlases. Each atlas is warped to fit a subject's
-neuroanatomy using nonlinear registration
-techniques \citep{Collins1995,Klein2009}.  Atlas labels are then transformed by
-this warping and a {\it label fusion} technique, such as voxel-wise voting, is
-used to merge the competing label definitions from each atlas into a final
-segmentation for a subject.
+Automated segmentation techniques overcome the need for human expertise
+by performing segmentations computationally. A popular class of
+automated methods, *multi-atlas-based segmentation*, rely
+on a set of expertly labeled neuroanatomical atlases. Each atlas is
+warped to fit a subject’s neuroanatomy using nonlinear registration
+techniques [@Collins1995; @Klein2009]. Atlas labels are then transformed
+by this warping and a *label fusion* technique, such as
+voxel-wise voting, is used to merge the competing label definitions from
+each atlas into a final segmentation for a subject.
 
-Many descriptions of multi-atlas-based segmentation algorithms report relying
-on an atlas library containing anywhere between 30 and 80 expertly labeled
-brains \citep{Heckemann2011,Collins2010,Aljabar2009,Leung2010,Lotjonen2010}.  As
-noted, the production of an atlas library requires significant manual effort,
-and is limited since the choice of atlases or segmentation protocol may not
-reflect the underlying neuroanatomical variability of the population under
-study or be suited to answer the research questions at hand.
+Many descriptions of multi-atlas-based segmentation algorithms report
+relying on an atlas library containing anywhere between 30 and 80
+expertly labeled brains
+[@Heckemann2011; @Collins2010; @Aljabar2009; @Leung2010; @Lotjonen2010].
+As noted, the production of an atlas library requires significant manual
+effort, and is limited since the choice of atlases or segmentation
+protocol may not reflect the underlying neuroanatomical variability of
+the population under study or be suited to answer the research questions
+at hand.
 
-In this paper we propose an automated segmentation method to address the above
-issues of existing multi-atlas-based methods. Principally, our method aims to
-dramatically reduce the number of manually labelled atlases needed (under 10).
-This is achieved by using the small atlas library to boot-strap a much larger
-"template library", which is then used to segment the subjects in a similar
-fashion to basic multi-atlas segmentation. This approach has the additional
-advantage of using the unique subject population on hand to initialize the
-segmentation process and improve accuracy.
+In this paper we propose an automated segmentation method to address the
+above issues of existing multi-atlas-based methods. Principally, our
+method aims to dramatically reduce the number of manually labelled
+atlases needed (under 10). This is achieved by using the small atlas
+library to boot-strap a much larger “template library”, which is then
+used to segment the subjects in a similar fashion to basic multi-atlas
+segmentation. This approach has the additional advantage of using the
+unique subject population on hand to initialize the segmentation process
+and improve accuracy.
 
-The essential insight of generating a template library is not new.  Heckemann
-\citep{Heckemann2006} compared generating a template library from a single atlas
-to standard multi-atlas segmentation and found poor performance and so deemed
-the approach as inviable.  The LEAP algorithm \citep{Wolz2010} proceeds by
-iteratively segmenting the unlabelled image most similar to the atlas library
-images and then incorporating the now-labelled image into the atlas library,
-but requires 30 starting atlases.  The novelty of our method is to demonstrate
-the possibility of producing comparable segmentation accuracy to these and
-other multi-atlas-based methods while using significantly fewer manually
+The essential insight of generating a template library is not new.
+Heckemann [@Heckemann2006] compared generating a template library from a
+single atlas to standard multi-atlas segmentation and found poor
+performance and so deemed the approach as inviable. The LEAP algorithm
+[@Wolz2010] proceeds by iteratively segmenting the unlabelled image most
+similar to the atlas library images and then incorporating the
+now-labelled image into the atlas library, but requires 30 starting
+atlases. The novelty of our method is to demonstrate the possibility of
+producing comparable segmentation accuracy to these and other
+multi-atlas-based methods while using significantly fewer manually
 created atlases.
 
-In our previous work \citep{Chakravarty2012}, we applied MAGeT brain to
-segmentation of the human striatum, globus pallidus, and thalamus using a
-single histologically-derived atlas. The main contribution of this paper is to
-extend our approach to the human hippocampus and perform a thorough validation
-over a range of atlas and template library sizes, which was not done in our
-previous work.  Due to the small number of atlases required, our method can
-easily accommodate different hippocampal definitions. Our aim is not to improve
-on segmentation accuracy beyond existing methods, but instead to provide a
-method that trades off manual segmentation expertise for computational
-processing time while providing sufficient accuracy for clinical and research
+In our previous work [@Chakravarty2012], we applied MAGeT brain to
+segmentation of the human striatum, globus pallidus, and thalamus using
+a single histologically-derived atlas. The main contribution of this
+paper is to extend our approach to the human hippocampus and perform a
+thorough validation over a range of atlas and template library sizes,
+which was not done in our previous work. Due to the small number of
+atlases required, our method can easily accommodate different
+hippocampal definitions. Our aim is not to improve on segmentation
+accuracy beyond existing methods, but instead to provide a method that
+trades off manual segmentation expertise for computational processing
+time while providing sufficient accuracy for clinical and research
 applications.
 
-\section{Materials and Methods} 
-\subsection{The Multiple Automatically Generated Templates (MAGeT) Algorithm}
+Materials and Methods
+=====================
 
-In this paper, we use the term {\it atlas} to mean any manually segmented MR
-image, and the term {\it atlas library} to mean a set of such images.  We use
-the term {\it template} to refer to any MR image, and associated labelling,
-used to segment another image, and the term {\it template library} to refer to
-a set of such images.  An atlas library may be used as a template library but,
-as we will discuss, a template library may also be composed of images with
+The Multiple Automatically Generated Templates (MAGeT) Algorithm
+----------------------------------------------------------------
+
+In this paper, we use the term *atlas* to mean any manually
+segmented MR image, and the term *atlas library* to mean a
+set of such images. We use the term *template* to refer to
+any MR image, and associated labelling, used to segment another image,
+and the term *template library* to refer to a set of such
+images. An atlas library may be used as a template library but, as we
+will discuss, a template library may also be composed of images with
 computer generated labellings.
 
-The segmentation approach we propose is best understood as an extension of
-basic multi-atlas segmentation \citep{Collins2010}.  In multi-atlas segmentation,
-an atlas library and unlabelled MR images are given as input.  Every atlas
-image is nonlinearly registered to each unlabelled image, and then each atlas'
-labels are propagated via the resulting transformations.  These labels are then
-fused to produce a single, definitive segmentation by some label fusion method
-(e.g.  voxel-wise majority vote).
+The segmentation approach we propose is best understood as an extension
+of basic multi-atlas segmentation [@Collins2010]. In multi-atlas
+segmentation, an atlas library and unlabelled MR images are given as
+input. Every atlas image is nonlinearly registered to each unlabelled
+image, and then each atlas’ labels are propagated via the resulting
+transformations. These labels are then fused to produce a single,
+definitive segmentation by some label fusion method (e.g. voxel-wise
+majority vote).
 
 Our extension adds a preliminary stage in which a template library is
-constructed from input images, and used in place of an atlas library in the
-standard multi-atlas-based method.  To create the template library, labels from
-each atlas image are propagated to each template library image via the
-transformation resulting from a non-linear registration between pair of images.
-As a result, each template library image has a label from each atlas.
-Basic multi-atlas segmentation is then used to produce segmentations for the
-entire set of unlabelled images (including those images used in the template
-library).
+constructed from input images, and used in place of an atlas library in
+the standard multi-atlas-based method. To create the template library,
+labels from each atlas image are propagated to each template library
+image via the transformation resulting from a non-linear registration
+between pair of images. As a result, each template library image has a
+label from each atlas. Basic multi-atlas segmentation is then used to
+produce segmentations for the entire set of unlabelled images (including
+those images used in the template library).
 
-Label fusion is performed by cross-correlation weighted voting, a strategy
-weighted towards an optimal combination of subjects from the template library
-which has been previously shown to improve segmentation
-accuracy \citep{Aljabar2009,Collins2010}. In this method, each template library
-image is ranked in similarity to each unlabelled image by the normalized
-cross-correlation of image intensities after linear registration in a region of
-interest (ROI) generously encompassing the hippocampus.  Only the top ranked
-template library image labels are used in a voxel-wise majority vote. The ROI
-is heuristically defined as the extent of all atlas labels after linear
-registration to the template, dilated by three voxels \citep{Chakravarty2012}.
+Label fusion is performed by cross-correlation weighted voting, a
+strategy weighted towards an optimal combination of subjects from the
+template library which has been previously shown to improve segmentation
+accuracy [@Aljabar2009; @Collins2010]. In this method, each template
+library image is ranked in similarity to each unlabelled image by the
+normalized cross-correlation of image intensities after linear
+registration in a region of interest (ROI) generously encompassing the
+hippocampus. Only the top ranked template library image labels are used
+in a voxel-wise majority vote. The ROI is heuristically defined as the
+extent of all atlas labels after linear registration to the template,
+dilated by three voxels [@Chakravarty2012].
 
-\subsection{MRI dataset evaluated}
+MRI dataset evaluated
+---------------------
 
-For evaluation purposes we used the publicly available IBSR dataset.  This
-dataset consists of T1-weighted MR image volumes from 18 subjects (4 females,
-14 males) with ages between 7 and 71 years. Image dimensions for all MR volumes
-are normalized to  $256  \times 256 \times 128$ voxels, with the voxel size
-ranging from $0.8 \times 0.8 \times 1.5 mm$ to $1.0 \times 1.0 \times 1.5 mm$.
-The images come 'positionally normalized' into the Talairach orientation
-(rotation only), and processed by the CMA 'autoseg' biasfield correction
-routines. The MR brain data sets and their manual segmentations are publicly
-available and were provided by the Center for Morphometric Analysis at
+For evaluation purposes we used the publicly available IBSR dataset.
+This dataset consists of T1-weighted MR image volumes from 18 subjects
+(4 females, 14 males) with ages between 7 and 71 years. Image dimensions
+for all MR volumes are normalized to $256  \times 256 \times 128$
+voxels, with the voxel size ranging from $0.8 \times 0.8 \times 1.5 mm$
+to $1.0 \times 1.0 \times 1.5 mm$. The images come ’positionally
+normalized’ into the Talairach orientation (rotation only), and
+processed by the CMA ’autoseg’ biasfield correction routines. The MR
+brain data sets and their manual segmentations are publicly available
+and were provided by the Center for Morphometric Analysis at
 Massachusetts General Hospital and are available at
-{\tt http://www.cma.mgh.harvard.edu/ibsr/}.
+http://www.cma.mgh.harvard.edu/ibsr/.
 
-\subsection{Image Processing and Registration Method}
-The N3 algorithm \citep{Sled1998} is first used to minimize the intensity
-nonuniformity in each image.  Image registration is carried out in two phases.
-In the first, a 12-parameter linear transformation (3 translations, rotations,
-scales, shears) is estimated between images using an algorithm that maximizes
-the correlation between blurred MR intensities and gradient magnitude over the
-whole brain \citep{Collins}.  In the second phase, nonlinear registration is
-completed using the ANIMAL algorithm \citep{Collins1995}: an iterative procedure
-that estimates a 3D deformation field between two MR images. At first, large
-deformations are estimated using blurred version of the input data. These
-larger deformations are then input to subsequent steps where the fit is refined
-by estimating smaller deformations on data blurred with a Gaussian kernel with
-a smaller FWHM. The final transformation is a set of local translations defined
-on a bed of equally spaced nodes that were estimated through the optimization
-of the correlation coefficient. For the purposes of this work we used the
-regularization parameters optimized in Robbins et al. \citep{Robbins2004}. It
-should be noted that the MAGeT brain algorithm is not dependent on this, or
-any, particular choice of registration method \citep{Chakravarty2012}.
+Image Processing and Registration Method
+----------------------------------------
 
-\subsection{Experiments}
+The N3 algorithm [@Sled1998] is first used to minimize the intensity
+nonuniformity in each image. Image registration is carried out in two
+phases. In the first, a 12-parameter linear transformation (3
+translations, rotations, scales, shears) is estimated between images
+using an algorithm that maximizes the correlation between blurred MR
+intensities and gradient magnitude over the whole brain [@Collins]. In
+the second phase, nonlinear registration is completed using the ANIMAL
+algorithm [@Collins1995]: an iterative procedure that estimates a 3D
+deformation field between two MR images. At first, large deformations
+are estimated using blurred version of the input data. These larger
+deformations are then input to subsequent steps where the fit is refined
+by estimating smaller deformations on data blurred with a Gaussian
+kernel with a smaller FWHM. The final transformation is a set of local
+translations defined on a bed of equally spaced nodes that were
+estimated through the optimization of the correlation coefficient. For
+the purposes of this work we used the regularization parameters
+optimized in Robbins et al. [@Robbins2004]. It should be noted that the
+MAGeT brain algorithm is not dependent on this, or any, particular
+choice of registration method [@Chakravarty2012].
 
-We explored how varying the size of the atlas library and the template library
-effects labeling accuracy.  For each parameter setting we conducted 30 rounds
-of random subsampling cross-validation using the 18 manual segmented templates
-from the IBSR dataset as input. In each round, atlases were randomly chosen
-from the IBSR dataset and the remaining images are used both as template
-library images and unlabeled subjects to be labeled using the MAGeT brain
-algorithm.  We varied the size of the atlas library from 3 to 8, and used
-cross-correlation weighted label fusion to select the top $n$ candidate
-templates from the remaining images.  $n$ was varied in the range $[3, 18-a]$,
-where $a$ is the size of the atlas library.
+Experiments
+-----------
 
-\subsection{Evaluation}
-\subsubsection{Goodness-of-fit}
+We explored how varying the size of the atlas library and the template
+library effects labeling accuracy. For each parameter setting we
+conducted 30 rounds of random subsampling cross-validation using the 18
+manual segmented templates from the IBSR dataset as input. In each
+round, atlases were randomly chosen from the IBSR dataset and the
+remaining images are used both as template library images and unlabeled
+subjects to be labeled using the MAGeT brain algorithm. We varied the
+size of the atlas library from 3 to 8, and used cross-correlation
+weighted label fusion to select the top $n$ candidate templates from the
+remaining images. $n$ was varied in the range $[3, 18-a]$, where $a$ is
+the size of the atlas library.
+
+Evaluation
+----------
+
+### Goodness-of-fit
+
 Automatically produced segmentations are evaluated against IBSR manual
-segmentations dataset using the Dice Kappa ($\kappa$) overlap metric, $\kappa =
+segmentations dataset using the Dice Kappa ($\kappa$) overlap metric,
+$\kappa =
 {2a}/{(2a+b+c)}$, where $a$ is the number of voxels common to both
-segmentations and $b+c$ is the sum of the voxels uniquely identified in either
-segmentation.
+segmentations and $b+c$ is the sum of the voxels uniquely identified in
+either segmentation.
 
-\subsubsection{Comparison Approaches}
-The resulting segmentations from each of our experiments are compared to those
-produced from two alternative segmentation approaches. The {\it single-atlas}
-approach uses one atlas to segment a unlabelled subject by directly propagating
-labels from the atlas by way of nonlinear registration.  We computed a
-single-atlas segmentation for each image in the IBSR dataset from each of the
-other 17 labelled images.  Similarly, we computed a segmentation for each image
-using the {\it basic multi-atlas} approach, described above, using the
-other 17 images as the atlas library.  Additionally, we also varied the number
-of atlas images used in the label fusion step by employing cross-correlation
-weighted voting.  In total we evaluated approximately $52,000$ segmentations
-for the work presented in this manuscript.
+### Comparison Approaches
 
-\section{Results}
+The resulting segmentations from each of our experiments are compared to
+those produced from two alternative segmentation approaches. The
+*single-atlas* approach uses one atlas to segment a
+unlabelled subject by directly propagating labels from the atlas by way
+of nonlinear registration. We computed a single-atlas segmentation for
+each image in the IBSR dataset from each of the other 17 labelled
+images. Similarly, we computed a segmentation for each image using the
+*basic multi-atlas* approach, described above, using the
+other 17 images as the atlas library. Additionally, we also varied the
+number of atlas images used in the label fusion step by employing
+cross-correlation weighted voting. In total we evaluated approximately
+$52,000$ segmentations for the work presented in this manuscript.
 
-Sample segmentations from a single IBSR subject compared with the gold-standard
-segmentation are in Fig.  \ref{montage}.  Qualitatively, as the size of the
-template library is increased, the number of false negatives in the hippocampal
-tail region is reduced, and segmentation accuracy increases.
+Results
+=======
 
-\begin{figure}[h]
-\begin{minipage}[b]{1.0\linewidth}
-  \centering
-  \includegraphics[width=\textwidth]{montage.png}
-\end{minipage}
-\caption{{\small
-Segmentations of a single subject are shown when three atlases are used, with
+Sample segmentations from a single IBSR subject compared with the
+gold-standard segmentation are in [Fig. 1](#fig:montage). Qualitatively, as the
+size of the template library is increased, the number of false negatives in the
+hippocampal tail region is reduced, and segmentation accuracy increases.
+
+![Segmentations of a single subject are shown when three atlases are used, with
 varying template library size $n$. Blue colouring represents agreement between
@@
 gold-standard labels), and red colouring indicates false negative voxels (i.e.
-voxels labelled in the gold-standard but not by MAGeT brain).
-}} 
-\label{montage}
-\end{figure}
+voxels labelled in the gold-standard but not by MAGeT brain).](montage.png){#fig:montage}
 
-MAGeT brain achieves a level of segmentation accuracy that is to within 2.0\%
-of the accuracy of the basic multi-atlas approach (Fig. \ref{results};
-mean $\kappa = 0.775$ vs mean $\kappa = 0.791$) in the best case. Importantly,
-to do this MAGeT brain only requires 8 manual segmented images whereas the
-basic multi-atlas requires 17 atlases.  This represents a significant
-savings in manual effort, and supports our contention that it is possible with
-MAGeT brain to trade a small decrease in accuracy for a significant decrease in
-the number of manual segmentations needed.
+MAGeT brain achieves a level of segmentation accuracy that is to within
+2.0% of the accuracy of the basic multi-atlas approach (Fig.
+*results*; mean $\kappa = 0.775$ vs mean $\kappa = 0.791$) in the best
+case. Importantly, to do this MAGeT brain only requires 8 manual
+segmented images whereas the basic multi-atlas requires 17 atlases. This
+represents a significant savings in manual effort, and supports our
+contention that it is possible with MAGeT brain to trade a small
+decrease in accuracy for a significant decrease in the number of manual
+segmentations needed.
 
@@
 
-\begin{figure}[t]
-%\begin{minipage}[b]{1.0\linewidth}
-  \centering
-  %\includegraphics[width=\textwidth]{results}
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\includegraphics[width=\maxwidth]{figure/performance} 
-\end{knitrout}
+```{r, echo=FALSE, fig.width=5, fig.cap="Mean performance of MAGeT brain with varying atlas library size and number of template labels fused.
+Also shown is the mean basic multi-atlas performance when using an atlas library of 17 images and varying the number of labels fused, as well as the mean performance of single-atlas segmentations.
+Data is fit with LOESS local regression smoothing.
+One standard deviation is shown in grey."}
+library(ggplot2)
+results = read.csv(file='data/results.csv')
+xcorr_vote= subset(results, method=='xcorr' | method == 'multiatlas_xcorr_vote' )
+qplot(x=top_n, y=k_hippocampus, data=xcorr_vote, colour=as.factor(num_atlases), geom=c('point'), alpha=I(0)) +
+  geom_smooth(method='loess')+ #line=1.5, formula=y~poly(x,2), method='lm') + 
+  geom_hline(aes(yintercept = mean(subset(results, method=='naive')$k_hippocampus), linetype="dashed")) + 
+  theme(legend.position="none") + 
+  ylab('Mean similarity score (Kappa)') + 
+  xlab('Number of template labels fused') +
+  annotate("text", 14.3, 0.712, label="Single-atlas", size=4, hjust=0)+
+  annotate("text", 15.3, 0.754, label="3 atlases", size=4, hjust=0)+
+  annotate("text", 14.4, 0.766, label="4 atlases", size=4, hjust=0)+
+  annotate("text", 13.5, 0.770, label="5 atlases", size=4, hjust=0)+
+  annotate("text", 12.5, 0.773, label="6 atlases", size=4, hjust=0)+
+  annotate("text", 11.5, 0.776, label="7 atlases", size=4, hjust=0)+
+  annotate("text", 10.3, 0.779, label="8 atlases", size=4, hjust=0)+
+  annotate("text", 13.5, 0.793, label="Basic multi-atlas", size=4, hjust=0)+
+  annotate("text", 14.6, 0.787, label="17 atlases", size=4, hjust=0)+
+  scale_colour_manual(values=c(
+    "#D73027","#FC8D59","#FEE08B","#D9EF8B", "#91CF60","#1A9850", "#000000")) + 
+  scale_linetype_manual(values=c(
+    "dotted", "solid", "solid", "solid",
+         "solid", "solid", "solid", "dashed")) +
+  coord_cartesian(ylim=c(0.703,0.803), xlim=c(3,18)) + scale_y_continuous(breaks=seq(0,1,0.01))   
+```
 
-%\end{minipage}
-\caption{{\small
-Mean performance of MAGeT brain with varying atlas library size and number of
-template labels fused. Also shown is the mean basic multi-atlas
-performance when using an atlas library of 17 images and varying the number of
-labels fused, as well as the mean performance of single-atlas segmentations.
-Data is fit with LOESS local regression smoothing. One standard deviation is
-shown in grey.
-}}
-\label{results}
-\end{figure}
+Discussion
+==========
 
-\section{Discussion}
+In this paper, we have demonstrated that accurate segmentations can be
+produced by automatically deriving a template library from a small set
+of input atlases. MAGeT brain segmentations were compared to both
+single-atlas segmentations and multi-atlas segmentations with
+cross-correlation voting. For the IBSR dataset, on average MAGeT brain
+achieves within 2.0% of the segmentation accuracy of the multi-atlas
+approach but requires only 8 atlases as compared to using the entire
+IBSR manually segmented library of 17 atlases.
 
-In this paper, we have demonstrated that accurate segmentations can be produced
-by automatically deriving a template library from a small set of input atlases.
-MAGeT brain segmentations were compared to both single-atlas segmentations and
-multi-atlas segmentations with cross-correlation voting.  For the IBSR dataset,
-on average MAGeT brain achieves within 2.0\% of the segmentation accuracy of
-the multi-atlas approach but requires only 8 atlases as compared to using the
-entire IBSR manually segmented library of 17 atlases.
+Lötjönen et al. [@Lotjonen2010] report a Kappa of 0.814 on hippocampal
+segmentation in the IBSR dataset using a multi-atlas approach where
+atlas selection is based on the similarity between atlas and unlabeled
+subject after nonlinear registration, and using STAPLE [@Warfield2004]
+for label fusion. Discrepancies between our results and the above
+results may be due to choice of registration algorithm, regularization
+parameters, or similarity metric for label fusion. Performance may also
+be affected by the variability in the IBSR data set (as previously noted
+by [@Klein2009]). Future work from our group will attempt to address
+some of these issues.
 
-L\"{o}tj\"{o}nen et al. \citep{Lotjonen2010} report a Kappa of 0.814 on
-hippocampal segmentation in the IBSR dataset using a multi-atlas approach where
-atlas selection is based on the similarity between atlas and unlabeled subject
-after nonlinear registration, and using STAPLE \citep{Warfield2004} for label
-fusion.  Discrepancies between our results and the above results may be due to
-choice of registration algorithm, regularization parameters, or similarity
-metric for label fusion.  Performance may also be affected by the variability
-in the IBSR data set (as previously noted by \citep{Klein2009}).  Future work
-from our group will attempt to address some of these issues.
+Acknowledgements
+================
 
-\section{Acknowledgements}
 Computations were performed on the gpc supercomputer at the SciNet HPC
-Consortium \citep{Loken2010}. SciNet is funded by: the Canada Foundation for
-Innovation under the auspices of Compute Canada; the Government of Ontario;
-Ontario Research Fund - Research Excellence; and the University of Toronto.
+Consortium [@Loken2010]. SciNet is funded by: the Canada Foundation for
+Innovation under the auspices of Compute Canada; the Government of
+Ontario; Ontario Research Fund - Research Excellence; and the University
+of Toronto.
 
-This work was supported by the Canadian Institutes of Health Research (CIHR),
-National Alliance for Research on Schizophrenia and Depression (NARSAD),
-Ontario Mental Health Foundation (OMHF), and the CAMH Foundation (Koerner New
-Scientist Program and Paul Garfinkel New Investigator Catalyst Fund).
-\bibliographystyle{IEEEbib}
-\bibliography{references}
-\end{document}
+This work was supported by the Canadian Institutes of Health Research
+(CIHR), National Alliance for Research on Schizophrenia and Depression
+(NARSAD), Ontario Mental Health Foundation (OMHF), and the CAMH
+Foundation (Koerner New Scientist Program and Paul Garfinkel New
+Investigator Catalyst Fund).

### 2016-05-26 paper.Rnw
Clean up of the Rnw source

--- 2012-11-08 paper.Rnw
@@
 \SweaveOpts{concordance=TRUE}
+
+<<setup include=FALSE, echo=FALSE>>=
+library(knitr)
+opts_chunk$set(
+concordance=TRUE
+)
+@
+
 \maketitle
@@
 \begin{figure}[t]
-%\begin{minipage}[b]{1.0\linewidth}
   \centering
-  %\includegraphics[width=\textwidth]{results}
-<<performance, fig=TRUE, cached=TRUE, echo=FALSE, warnings=FALSE >>=
+
+<<performance, cached=TRUE, echo=FALSE, warnings=FALSE >>=
 library(ggplot2)
+
 results = read.csv(file='data/results.csv')
-#results = read.csv(file='data/results_with_9multiatlas.csv')
 xcorr_vote= subset(results, method=='xcorr' | method == 'multiatlas_xcorr_vote' )
+
 qplot(x=top_n, y=k_hippocampus, data=xcorr_vote, colour=as.factor(num_atlases), geom=c('point'), alpha=I(0)) +
@@
 @
-%\end{minipage}
 \caption{{\small
